---
title: "08_17_2020_HW"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rethinking)
library(brms)
library(tidyverse)
```

# Assignment

Revisit the following homework problems and try to fit the with brms.  Make your first attempt without looking at the rethinking to brms translation, but if you get stuck definitely look!  Compare the coefficients or predictions that you obtain with brms and those with quap or ulam.

## 1. 4H1, 4H2 (you probably need the function `posterior_predict()`)

## Hard.

### 4H1. The weights listed below were recorded in the !Kung census, but heights were not recorded for these individuals. Provide predicted heights and 89% intervals (either HPDI or PI) for each of these individuals. That is, fill in the table below, using model-based predictions. Individual weight expected height 89% interval


```{r}
data(Howell1)
dat <- Howell1[Howell1$age >= 18,]
weights <- c(46.95,43.72,64.78,32.59,54.63)
H1 <- quap(
alist(
height ~ dnorm( mu , sigma ) ,
mu <- a + b*( weight - mean(weight) ) ,
a ~ dnorm( 178 , 20 ) ,
b ~ dlnorm( 0 , 1 ) ,
sigma ~ dunif( 0 , 50 )
) , data=dat)

heights <- sim(H1,
               data = data.frame(weight=weights),
               n=1e3)
#head(heights)
apply( heights , 2 , mean)
apply( heights , 2 , HPDI , prob=0.89 )
```

|Individual|Weight|Expected Height| 89% Interval|
|-----|-----|-----|-------------|
|1|46.95|135.88|121.76-152.03|
|2|43.72|129.52|112.76-142.83|
|3|64.78|167.02|151.08-180.49|
|4|32.59|110.19|95.12-125.39|
|5|54.63|149.1716|134.16-165.24|

### brms answers

```{r}
dat$weight2 <- dat$weight - mean(dat$weight)
weights <- data.frame(weight2 = c(46.95,43.72,64.78,32.59,54.63) - mean(dat$weight))
m4.1_brm <- brm(height ~ weight2, data=dat,
               prior=c(
                 set_prior("normal(178,20)", class = "Intercept"),
                 set_prior("lognormal(0,1)", class = "b",lb=0),
                 set_prior("exponential(1)", class = "sigma"))
               )
m4.1_brm

# Predict
preds <- posterior_predict(m4.1_brm, newdata = weights)
head(preds)

apply(preds,2, mean)
apply(preds,2, HPDI)

```

### Rethinking answers

|Individual|Weight|Expected Height| 89% Interval|
|-----|-----|-----|-------------|
|1|46.95|135.88|121.76-152.03|
|2|43.72|129.52|112.76-142.83|
|3|64.78|167.02|151.08-180.49|
|4|32.59|110.19|95.12-125.39|
|5|54.63|149.1716|134.16-165.24|

### brms answers

|Individual|Weight|Expected Height| 89% Interval|
|-----|-----|-----|-------------|
|1|46.95|156.5332|148.8583-165.0345|
|2|43.72|153.4227|145.2876-161.2894|
|3|64.78|172.5122|164.4680-181.0276|
|4|32.59|143.4289|135.4683-151.8551|
|5|54.63|163.2118|154.9893-171.2563|


### 4H2. Select out all the rows in the Howell1 data with ages below 18 years of age. If you do it right, you should end up with a new data frame with 192 rows in it.

```{r}
young.dat <- Howell1[Howell1$age < 18, ]
```


(a) Fit a linear regression to these data, using quap. Present and interpret the estimates. For
  every 10 units of increase in weight, how much taller does the model predict a child gets?
  
```{r}
plot(young.dat$height ~ young.dat$weight)
xbar <- mean(young.dat$weight)
H2 <- quap(alist(
  height ~ dnorm(mu , sigma) ,
  mu <- a + b * (weight - xbar) ,
  a ~ dnorm(178 , 20) ,
  b ~ dlnorm(0 , 1) ,
  sigma ~ dunif(0 , 50)
) ,
data = young.dat)
precis(H2)
pairs(H2)
2.72 * 10 # 27.2 cm increase per 10 kg
```

```{r}
young.dat$weight2 <- young.dat$weight - xbar
mean(young.dat$weight)
m4.2_brm <- brm(height ~ weight2, data=dat,
               prior=c(
                 set_prior("normal(0,5)", class = "Intercept"),
                 set_prior("lognormal(0,1)", class = "b",lb=0),
                 set_prior("exponential(1)", class = "sigma"))
               )
m4.2_brm
pairs(m4.2_brm)
0.90 * 10 #9 cm per 10 kg?
```


  (b) Plot the raw data, with height on the vertical axis and weight on the horizontal axis. Superimpose the MAP regression line and 89% HPDI for the mean. Also superimpose the 89% HPDI for predicted heights.
  
```{r}
post <- extract.samples(H2)
weight.seq <- seq.int(from=min(young.dat$weight), to=max(young.dat$weight), by=1)
mu <- link(H2, data=data.frame(weight=weight.seq))
mu.mean <- apply( mu , 2 , mean )
mu.HPDI <- apply( mu , 2 , HPDI , prob=0.89 )
sim <- sim(H2, data=data.frame(weight=weight.seq))
sim.HPDI <- apply(sim, 2, HPDI, prob=0.89)
plot( height ~ weight , data=young.dat , col=rangi2 )
# plot the MAP line, aka the mean mu for each weight
lines( weight.seq , mu.mean )
shade( sim.HPDI , weight.seq)
# plot a shaded region for 89% HPDI
shade( mu.HPDI , weight.seq)
```

```{r}
weight.seq <- seq.int(from=min(young.dat$weight2), to=max(young.dat$weight2), by=1)
post <- posterior_epred(m4.2_brm, newdata = data.frame(weight2 = weight.seq))
post.mean <- apply( post , 2 , mean )
post.HPDI <- apply( post , 2 , HPDI , prob=0.89 )

plot( height ~ weight , data=young.dat , col=rangi2 )
# plot the MAP line, aka the mean mu for each weight
lines( weight.seq , post.mean)

# plot a shaded region for 89% HPDI
shade(post.HPDI , weight.seq)
```


  (c) What aspects of the model fit concern you? Describe the kinds of assumptions you would change, if any, to improve the model. You don’t have to write any new code. Just explain what the model appears to be doing a bad job of, and what you hypothesize would be a better model
  
Model is very poor at predicting heights as you move to the extremes of the weight ranges.

```{r}
young.dat$weight_s <- ( young.dat$weight - mean(young.dat$weight) )/sd(young.dat$weight)
young.dat$weight_s2 <- young.dat$weight_s^2
young.dat$weight_s3 <- young.dat$weight_s^3
H2.2 <- quap(
alist(
height ~ dnorm( mu , sigma ) ,
mu <- a + b1*weight_s + b2*weight_s2 + b3*weight_s3 ,
a ~ dnorm( 178 , 20 ) ,
b1 ~ dlnorm( 0 , 1 ) ,
b2 ~ dnorm( 0 , 10 ) ,
b3 ~ dnorm( 0 , 10 ) ,
sigma ~ dunif( 0 , 50 )
) ,
data=young.dat )
weight.seq <- seq( from=-2.2 , to=3 , length.out=30 )
pred_dat <- list( weight_s=weight.seq , weight_s2=weight.seq^2, weight_s3=weight.seq^3 )
head(pred_dat)
mu <- link( H2.2 , data=pred_dat )
mu.mean <- apply( mu , 2 , mean )
mu.PI <- apply( mu , 2 , PI , prob=0.89 )
sim.height <- sim( H2.2 , data=pred_dat )
height.PI <- apply( sim.height , 2 , PI , prob=0.89 )
plot( height ~ weight_s , young.dat , col=col.alpha(rangi2,0.5) )
lines( weight.seq , mu.mean )
shade( mu.PI , weight.seq )
shade( height.PI , weight.seq )
```


* From chapter 9: 8M1 (remember that the problem numbers were offset it is actually called 9M1 in the Nov 24 PDF)