---
title: "04_03_2020_Notes"
author: "John Davis"
date: "04_03_2020"
output:
  github_document:
    pandoc_args: --webtex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(rethinking)
library(tidyverse)
```

## 14.4 Social relations as correlated varying effects

```{r}
data(KosterLeckie)
```

```{r}
plot(kl_dyads$giftsAB,kl_dyads$giftsBA)
abline(a=0,b=1)
cor(kl_dyads$giftsAB, kl_dyads$giftsBA)
```

```{r}
kl_data <- list(
  N = nrow(kl_dyads),
  N_households = max(kl_dyads$hidB),
  did = kl_dyads$did,
  hidA = kl_dyads$hidA,
  hidB = kl_dyads$hidB,
  giftsAB = kl_dyads$giftsAB,
  giftsBA = kl_dyads$giftsBA
)

m14.4 <- ulam(
  alist(
    giftsAB ~ poisson( lambdaAB ), # A to B model
    giftsBA ~ poisson( lambdaBA ), # B to A model
    # A to B, give/receive coefficient + The dyad specific rate A to B
    log(lambdaAB) <- a + gr[hidA,1] + gr[hidB,2] + d[did,1],
    # B to A, give/receive coefficient + The dyad specific rate B to A
    log(lambdaBA) <- a + gr[hidB,1] + gr[hidA,2] + d[did,2],
    a ~ normal(0,1), # Average gift giving across all dyads
    
    ## gr matrix of varying effects
    # vector of 2, a give and receive for each house (25,2)
    vector[2]:gr[N_households] ~ multi_normal(0,Rho_gr,sigma_gr),
    # Expected almost no correlation between give and receive
    Rho_gr ~ lkj_corr(4),
    # variation
    sigma_gr ~ exponential(1),
    
    ## dyad effects
    # Noncentered parametized, sigma repeated twice due to same value in
    # both directions. Finding dyad specific effects, mirrored/doubled due
    # to being the same sigma. (300,2) 
    transpars> matrix[N,2]:d <- compose_noncentered( rep_vector(sigma_d,2) , L_Rho_d , z ),
    # matrix for z in noncentered model
    matrix[2,N]:z ~ normal( 0 , 1 ),
    # Almost no correlation prior
    cholesky_factor_corr[2]:L_Rho_d ~ lkj_corr_cholesky( 8 ),
    # shared sigma_d
    sigma_d ~ exponential(1),
    ## compute correlation matrix for dyads
    gq> matrix[2,2]:Rho_d <<- Chol_to_Corr( L_Rho_d )
    ),
  data = kl_data, chains=4, cores=4, iter=2000
  )
```

```{r}
precis(m14.4, depth=3, pars=c("Rho_gr","sigma_gr"))
```
`Rho_gr` Correlation between gift giving and receiving is generally negative, across all dyads if you give more, you receive less  
`sigma_gr[1]` = giving and `sigma_gr[2]` = receiving. Rates of giving are more variable than receiving.

```{r}
post <- extract.samples( m14.4 )
g <- sapply( 1:25 , function(i) post$a + post$gr[,i,1] )
r <- sapply( 1:25 , function(i) post$a + post$gr[,i,2] )
Eg_mu <- apply( exp(g) , 2 , mean )
Er_mu <- apply( exp(r) , 2 , mean )
```

```{r}
plot(exp(g[,1]),exp(r[,1]))
```

```{r}
plot(
  NULL ,
  xlim = c(0, 8.6) ,
  ylim = c(0, 8.6) ,
  xlab = "generalized giving" ,
  ylab = "generalized receiving" ,
  lwd = 1.5
)
abline(a = 0, b = 1, lty = 2)
# ellipses
library(ellipse)
for (i in 1:25) {
  Sigma <- cov(cbind(g[, i] , r[, i]))
  Mu <- c(mean(g[, i]) , mean(r[, i]))
  for (l in c(0.5)) {
    el <- ellipse(Sigma , centre = Mu , level = l)
    lines(exp(el) , col = col.alpha("black", 0.5))
  }
}
# household means
points(Eg_mu ,
       Er_mu ,
       pch = 21 ,
       bg = "white" ,
       lwd = 1.5)
```

```{r}
precis( m14.4 , depth=3 , pars=c("Rho_d","sigma_d") )
```
Families that give to one family also receive from the same family more

```{r}
dy1 <- apply( post$d[,,1] , 2 , mean )
dy2 <- apply( post$d[,,2] , 2 , mean )
plot( dy1 , dy2 )
```

## 14.5. Continuous categories and the Gaussian process
### 14.5.1. Example: Spatial autocorrelation in Oceanic tools

```{r}
# load the distance matrix
data(islandsDistMatrix)
# display (measured in thousands of km)
Dmat <- islandsDistMatrix
colnames(Dmat) <- c("Ml","Ti","SC","Ya","Fi","Tr","Ch","Mn","To","Ha")
round(Dmat,1)
```

```{r}
# linear
curve(exp(-1*x), from=0, to=4, lty=2,
       xlab="distance", ylab="correlation")
# squared
curve( exp(-1*x^2) , add=TRUE )
```

```{r}
data(Kline2) # load the ordinary data, now with coordinates
d <- Kline2
d$society <- 1:10 # index observations

dat_list <- list(
  T = d$total_tools,
  P = d$population,
  society = d$society,
  Dmat = islandsDistMatrix
)

m14.7 <- ulam(
  alist(
    T ~ dpois(lambda),
    lambda <- (a * P ^ b / g) * exp(k[society]),
    vector[10]:k ~ multi_normal(0, SIGMA),
    matrix[10, 10]:SIGMA <- cov_GPL2(Dmat, etasq, rhosq, 0.01),
    c(a, b, g) ~ dexp(1),
    etasq ~ dexp(2),
    rhosq ~ dexp(0.5)
  ),
  data = dat_list,
  chains = 4,
  cores = 4,
  iter = 2000
)
```

```{r}
precis( m14.7 , depth=3 )
```

```{r}
post <- extract.samples(m14.7)
# plot the posterior median covariance function
plot(
  NULL ,
  xlab = "distance (thousand km)" ,
  ylab = "covariance" ,
  xlim = c(0, 10) ,
  ylim = c(0, 2)
)
# compute posterior mean covariance
x_seq <- seq(from = 0 ,
             to = 10 ,
             length.out = 100)
pmcov <-
  sapply(x_seq , function(x)
    post$etasq * exp(-post$rhosq * x ^ 2))
pmcov_mu <- apply(pmcov , 2 , mean)
lines(x_seq , pmcov_mu , lwd = 2)
# plot 60 functions sampled from posterior
for (i in 1:50)
  curve(post$etasq[i] * exp(-post$rhosq[i] * x ^ 2) ,
        add = TRUE ,
        col = col.alpha("black", 0.3))

```

```{r}
# compute posterior median covariance among societies
K <- matrix(0, nrow = 10, ncol = 10)
for (i in 1:10)
  for (j in 1:10)
    K[i, j] <- median(post$etasq) *
  exp(-median(post$rhosq) * islandsDistMatrix[i, j] ^ 2)
diag(K) <- median(post$etasq) + 0.01
K
```

```{r}
# convert to correlation matrix
Rho <- round( cov2cor(K) , 2 )
# add row/col names for convenience
colnames(Rho) <- c("Ml","Ti","SC","Ya","Fi","Tr","Ch","Mn","To","Ha")
rownames(Rho) <- colnames(Rho)
Rho
```

```{r}
# scale point size to logpop
psize <- d$logpop / max(d$logpop)
psize <- exp(psize * 1.5) - 2
# plot raw data and labels
plot(
  d$lon2 ,
  d$lat ,
  xlab = "longitude" ,
  ylab = "latitude" ,
  col = rangi2 ,
  cex = psize ,
  pch = 16 ,
  xlim = c(-50, 30)
)
labels <- as.character(d$culture)
text(
  d$lon2 ,
  d$lat ,
  labels = labels ,
  cex = 0.7 ,
  pos = c(2, 4, 3, 3, 4, 1, 3, 2, 4, 2)
)
# overlay lines shaded by Rho
for (i in 1:10)
  for (j in 1:10)
    if (i < j)
      lines(
        c(d$lon2[i], d$lon2[j]) ,
        c(d$lat[i], d$lat[j]) ,
        lwd = 2 ,
        col = col.alpha("black", Rho[i, j] ^ 2)
      )
```

```{r}
# compute posterior median relationship, ignoring distance
logpop.seq <- seq(from = 6 ,
                  to = 14 ,
                  length.out = 30)
lambda <-
  sapply(logpop.seq , function(lp)
    exp(post$a + post$b * lp))
lambda.median <- apply(lambda , 2 , median)
lambda.PI80 <- apply(lambda , 2 , PI , prob = 0.8)
# plot raw data and labels
plot(
  d$logpop ,
  d$total_tools ,
  col = rangi2 ,
  cex = psize ,
  pch = 16 ,
  xlab = "log population" ,
  ylab = "total tools"
)
text(
  d$logpop ,
  d$total_tools ,
  labels = labels ,
  cex = 0.7 ,
  pos = c(4, 3, 4, 2, 2, 1, 4, 4, 4, 2)
)
# display posterior predictions
lines(logpop.seq , lambda.median , lty = 2)
lines(logpop.seq , lambda.PI80[1, ] , lty = 2)
lines(logpop.seq , lambda.PI80[2, ] , lty = 2)
# overlay correlations
for (i in 1:10)
  for (j in 1:10)
    if (i < j)
      lines(
        c(d$logpop[i], d$logpop[j]) ,
        c(d$total_tools[i], d$total_tools[j]) ,
        lwd = 2 ,
        col = col.alpha("black", Rho[i, j] ^ 2)
      )
```

```{r}
m14.7nc <- ulam(
  alist(
    T ~ dpois(lambda),
    lambda <- (a * P ^ b / g) * exp(k[society]),
    # non-centered Gaussian Process prior
    transpars > vector[10]:k <<- L_SIGMA * z,
    vector[10]:z ~ normal(0 , 1),
    transpars > matrix[10, 10]:L_SIGMA <<- cholesky_decompose(SIGMA),
    transpars > matrix[10, 10]:SIGMA <-
      cov_GPL2(Dmat , etasq , rhosq , 0.01),
    c(a, b, g) ~ dexp(1),
    etasq ~ dexp(2),
    rhosq ~ dexp(0.5)
  ),
  data = dat_list ,
  chains = 4 ,
  cores = 4 ,
  iter = 2000
)
```

### 14.5.2. Example: Phylogenetic distance

```{r}
data(Primates301)
data(Primates301_nex)
# plot it using ape package - install.packages('ape') if needed
library(ape)
plot(
  ladderize(Primates301_nex) ,
  type = "fan" ,
  font = 1 ,
  no.margin = TRUE ,
  label.offset = 1 ,
  cex = 0.5
)
```

```{r}
d <- Primates301
d$name <- as.character(d$name)
dstan <- d[ complete.cases( d$group_size , d$body , d$brain ) , ]
spp_obs <- dstan$name
```

```{r}
dat_list <- list(
  N_spp = nrow(dstan),
  M = standardize(log(dstan$body)),
  B = standardize(log(dstan$brain)),
  G = standardize(log(dstan$group_size)),
  Imat = diag(nrow(dstan))
)

m14.8 <- ulam(
  alist(
    B ~ multi_normal(mu , SIGMA),
    mu <- a + bM * M + bG * G,
    matrix[N_spp, N_spp]:SIGMA <- Imat * sigma_sq,
    a ~ normal(0 , 1),
    c(bM, bG) ~ normal(0 , 0.5),
    sigma_sq ~ exponential(1)
  ),
  data = dat_list ,
  chains = 4 ,
  cores = 4
)

precis(m14.8)
```

```{r}
tree_trimmed <- keep.tip( Primates301_nex, spp_obs )
Rbm <- corBrownian( phy=tree_trimmed )
V <- vcv(Rbm)
Dmat <- cophenetic( tree_trimmed )
plot( Dmat , V , xlab="phylogenetic distance" , ylab="covariance" )
```

```{r}
image(V)
```

```{r}
image(Dmat)
```

```{r}
# put species in right order
dat_list$V <- V[spp_obs , spp_obs]

# convert to correlation matrix
dat_list$R <- dat_list$V / max(V)

# Brownian motion model
m14.9 <- ulam(
  alist(
    B ~ multi_normal(mu , SIGMA),
    mu <- a + bM * M + bG * G,
    matrix[N_spp, N_spp]:SIGMA <- R * sigma_sq,
    a ~ normal(0 , 1),
    c(bM, bG) ~ normal(0 , 0.5),
    sigma_sq ~ exponential(1)
  ),
  data = dat_list ,
  chains = 4 ,
  cores = 4
)

precis(m14.9)
```

```{r}
# add scaled and reordered distance matrix
dat_list$Dmat <- Dmat[spp_obs , spp_obs] / max(Dmat)

m14.10 <- ulam(
  alist(
    B ~ multi_normal(mu , SIGMA),
    mu <- a + bM * M + bG * G,
    matrix[N_spp, N_spp]:SIGMA <-
      cov_GPL1(Dmat , etasq , rhosq , 0.01),
    a ~ normal(0, 1),
    c(bM, bG) ~ normal(0, 0.5),
    etasq ~ half_normal(1, 0.25),
    rhosq ~ half_normal(3, 0.25)
  ),
  data = dat_list ,
  chains = 4 ,
  cores = 4
)

precis(m14.10)

```

```{r}
post <- extract.samples(m14.10)

plot(
  NULL ,
  xlim = c(0, max(dat_list$Dmat)) ,
  ylim = c(0, 1.5) ,
  xlab = "phylogenetic distance" ,
  ylab = "covariance"
)

# posterior
for (i in 1:30)
  curve(post$etasq[i] * exp(-post$rhosq[i] * x) ,
        add = TRUE ,
        col = rangi2)
# prior mean and 89% interval
eta <- abs(rnorm(1e3, 1, 0.25))
rho <- abs(rnorm(1e3, 3, 0.25))
d_seq <- seq(from = 0,
             to = 1,
             length.out = 50)
K <- sapply(d_seq , function(x)
  eta * exp(-rho * x))
lines(d_seq , colMeans(K) , lwd = 2)
shade(apply(K, 2, PI) , d_seq)
text(0.5 , 0.5 , "prior")
text(0.2 , 0.1 , "posterior" , col = rangi2)
```

