---
title: "01_31_2020_HW"
author: "John D."
date: "1/30/2020"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=F}
library(tidyverse)
library(rethinking)
```

Please find attached csv.  This csv has measurements of tomato internodes and petioles from wildtype (Moneymaker) and various phytochrome mutant lines.  Measurements were made at 3 time points, 21, 28, and 35 days after germination under two treatments, simulated sun and simulated shade.

for today let's focus on day 35.  Also let's focus on total stem length.  So: first

```{r}
dat <- read.csv("figure4phyE.csv")
head(dat)
```


## Q1)

  a) subset the data for day 35
  
```{r}
dat_35 <- dat %>%
  filter(day == 35)
```

  b) create a new column "stem_length" that is the sum of epi, int1, int2, and int3
  
```{r}
dat_35$stem_length <- dat_35$epi + dat_35$int1 + dat_35$int2 + dat_35$int3
```

  c) although flats are listed as 1-6, flats in sun and shade are separate. Create a new column "flat2" that corrects for this.

Ultimately you want to know if any of the mutants have a different length from Moneymaker, in sun or in shade, or if the response to shade differs.

```{r}
dat_35$flat2 <- paste(dat_35$treatment, dat_35$flat, sep = "_")

dat_35_list <- list(
  G = as.integer(dat_35$genotype),
  T = ifelse(dat_35$treatment == "sun", 0L, 1L),
  F = as.integer(as.factor(dat_35$flat2)),
  L = dat_35$stem_length
)

str(dat_35_list)

summary(dat_35)
```



## Q2) Fit 3 models, all of which include genotype and treatment

  a) do not include flat

```{r}
m_1 <- ulam(
  alist(
    L ~ dnorm(mu, sigma),
    mu <- a[G] + bT*T + bGT[G]*T,
    a[G] ~ dnorm(145, 50),
    bT ~ dnorm(0, 25),
    bGT[G] ~ dnorm(0, 25),
    sigma ~ dexp(1)
  ),
  data = dat_35_list ,
  chains = 4 ,
  log_lik = TRUE,
  cores = 4
)
```

  b) include flat without pooling

```{r}
m_2 <- ulam(
  alist(
    L ~ dnorm(mu, sigma),
    mu <- a[G] + bT*T + bGT[G]*T + bF[F],
    a[G] ~ dnorm(145, 50),
    bT ~ dnorm(0, 25),
    bGT[G] ~ dnorm(0, 25),
    bF[F] ~ dnorm(0,2),
    sigma ~ dexp(1)
  ),
  data = dat_35_list ,
  chains = 4 ,
  log_lik = TRUE,
  cores = 4
)
```

  c) use a hierarchical model that allows partial pooling across flats

```{r}
m_3 <- ulam(
  alist(
    L ~ dnorm(mu, sigma),
    mu <- a[G] + bT*T + bGT[G]*T + bF[F],
    a[G] ~ dnorm(145, 50),
    bT ~ dnorm(0, 25),
    bGT[G] ~ dnorm(0, 25),
    bF[F] ~ dnorm(f_bar, sigma_f),
    f_bar ~ dnorm(0,2),
    sigma_f ~ dexp(1),
    sigma ~ dexp(1)
  ),
  data = dat_35_list ,
  chains = 4 ,
  log_lik = TRUE,
  cores = 4
)
```

Should probably make a different model

```{r}
m_4 <- ulam(
  alist(
    L ~ dnorm(mu, sigma),
    mu <- a[G] + bT*T + bGT[G]*T + bF[F],
    a[G] ~ dnorm(a_bar, sigma_a),
    bT ~ dnorm(0, 25),
    bGT[G] ~ dnorm(0, 25),
    bF[F] ~ dnorm(0, sigma_f),
    a_bar ~ dnorm(145,15),
    c(sigma_a, sigma_f) ~ dexp(1),
    sigma ~ dexp(1)
  ),
  data = dat_35_list ,
  chains = 4 ,
  log_lik = TRUE,
  cores = 4,
  iter = 5000
)
```


## Q3) Compare the models, which is preferred?

```{r}
precis(m_1, depth = 2)
precis(m_2, depth = 2)
precis(m_3, depth = 2)
precis(m_4, depth = 2)
compare(m_1, m_2, m_3, m_4)
```

Model 3 is the best model based on WAIC, but it's pretty close. Error of m_4 is within WAIC of m_3

## Q4) Using the hierarchical model, make posterior predictions

  a) for average cluster
  
```{r}
post <- extract.samples(m_4)

link_abar <- function(treatment){
  pred <- with(post, a_bar + bT*treatment)
  return(pred)
}

raw <- sapply( 0:1 , function(i) link_abar( i ) )
mu <- apply( raw , 2 , mean )
ci <- apply( raw , 2 , PI )

plot( NULL , xlab="treatment" , ylab="stem length", ylim=c(0,300) , xaxt="n" , xlim=c(1,2) )
axis( 1 , at=1:2 , labels=c("Sun","Shade"))
lines( 1:2 , mu )
shade( ci , 1:2 )
```

  b) for same clusters

```{r}
link_same <- function(treatment, genotype = 1, flat = 1) {
  pred <- with(post, a[,genotype] + bF[,flat] + bT*treatment)
  return(pred)
}
raw <- sapply(0:1 , function(i) link_same( i , genotype = 1 , flat = 1))
mu <- apply( raw , 2 , mean )
ci <- apply( raw , 2 , PI )

plot( NULL , xlab="treatment" , ylab="stem length", ylim=c(0,300) , xaxt="n" , xlim=c(1,2), main = "Money Maker")
axis( 1 , at=1:2 , labels=c("Sun","Shade"))
lines( 1:2 , mu )
shade( ci , 1:2 )
```

  c) showing the "marginal" from cluster

```{r}
a_sim <- with( post , rnorm( length(post$a_bar) , a_bar , sigma_a ) )
link_asim <- function( treatment ) {
  pred <- with( post , a_sim + bT*treatment)
  return(pred)
}

raw <- sapply( 0:1 , function(i) link_asim( i ))
mu <- apply( raw , 2 , mean )
ci <- apply( raw , 2 , PI )

plot( NULL , xlab="treatment" , ylab="stem length", ylim=c(0,300) , xaxt="n" , xlim=c(1,2) )
axis( 1 , at=1:2 , labels=c("Sun","Shade"))
lines( 1:2 , mu )
shade( ci , 1:2 )
```

  d) showing new clusters.

```{r}
plot( NULL , xlab="treatment" , ylab="stem length", ylim=c(0,300) , xaxt="n" , xlim=c(1,2) )
axis( 1 , at=1:2 , labels=c("Sun","Shade"))
for ( i in 1:100 ) lines( 1:2 , raw[i,] , col=col.alpha("black",0.25) , lwd=2 )
```

## Q5) Reparameterize the model to help with divergent transitions (even if there aren't any)

```{r}
m_5 <- ulam(
  alist(
    L ~ dnorm(mu, sigma),
    mu <- a_bar + bG[G]*sigma_g + # genotype intercepts
          bT*T + # treatmet
          bGT[G]*sigma_g*T + # interaction
          bF[F]*sigma_f, # flat intercepts
    bG[G] ~ dnorm(0, 25),
    bT ~ dnorm(0, 25),
    bGT[G] ~ dnorm(0, 25),
    bF[F] ~ dnorm(0, 5),
    a_bar ~ dnorm(145,15),
    c(sigma_g, sigma_f) ~ dexp(1),
    sigma ~ dexp(1)
  ),
  data = dat_35_list ,
  chains = 4 ,
  log_lik = TRUE,
  cores = 4,
  iter = 5000
)

compare(m_1, m_2, m_3, m_4, m_5)
```

No more divergent transitions. Also samples fast

## Q6--optional)

  a) Which genotypes differ from MoneyMaker in Sun conditions?
  
```{r}
sun_sample_means <- post$a
colnames(sun_sample_means) <- levels(dat_35$genotype)
sun_sample_means <- as.data.frame(sun_sample_means) %>%
  gather(key = "Genotype", value = "Stem_length", 1:6)

ggplot(sun_sample_means, aes(x = Genotype, y = Stem_length)) +
  geom_boxplot()

plot(precis(m_5, depth = 2),
     pars = c("bG[1]","bG[2]","bG[3]","bG[4]","bG[5]","bG[6]"),
     labels = levels(dat_35$genotype),
     main = "Genotype in Sun"
)
```
  
  b) Which genotypes differ from MoneyMaker in Shade conditions?

```{r}
shade_sample_means <- post$a + post$bGT + matrix(rep(post$bT,6),ncol = 6)
colnames(shade_sample_means) <- levels(dat_35$genotype)
shade_sample_means <- as.data.frame(shade_sample_means) %>%
  gather(key = "Genotype", value = "Stem_length", 1:6)

ggplot(shade_sample_means, aes(x = Genotype, y = Stem_length)) +
  geom_boxplot()

plot(precis(m_5, depth = 2),
     pars = c("bGT[1]","bGT[2]","bGT[3]","bGT[4]","bGT[5]","bGT[6]"),
     labels = levels(dat_35$genotype),
     main = "Genotype by Treatment"
)
```

  c) Which genotypes differ from MoneyMaker in their response to shade (difference in sun vs shade)?
  
```{r}
sun_sample_means$Treatment <- "Sun"
shade_sample_means$Treatment <- "Shade"
sample_means <- rbind(sun_sample_means,shade_sample_means)

ggplot(sample_means, aes(x = Genotype, y = Stem_length, color = Treatment)) +
  geom_boxplot()

diffs <- data.frame(Genotype = sun_sample_means$Genotype,
                    Difference = shade_sample_means$Stem_length - sun_sample_means$Stem_length
)

diff_summaries <- diffs %>%
  group_by(Genotype) %>%
  group_map(~ summary(.x))

diff_summaries <- as.data.frame(matrix(unlist(diff_summaries),6,6))
colnames(diff_summaries) <- levels(dat_35$genotype)
diff_summaries

ggplot(diffs, aes(x = Genotype, y = Difference)) +
  geom_boxplot()
```

