---
title: "03_13_2020_HW"
author: "John Davis"
date: "03_13_2020"
output:
  github_document:
    pandoc_args: --webtex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(rethinking)
library(tidyverse)
```

## 14M3. Re-estimate the varying slopes model for the UCBadmit data, now using a non-centered parameterization. Compare the efficiency of the forms of the model, using n_eff. Which is better? Which chain sampled faster?

```{r}
data(UCBadmit)
d <- UCBadmit

dat_list <- list(
  admit = d$admit,
  applications = d$applications,
  gender = ifelse(d$applicant.gender == "male" , 1 , 2),
  dept = rep(1:6, each = 2)
)

system.time(
  m14m3.centered <- ulam(
    alist(
      admit ~ dbinom(applications, p),
      logit(p) <- a[dept] + beta[gender, dept],
      # adaptive prior
      vector[6]:beta[gender] ~ multi_normal(0, Rho_gender, sigma_gender),
      # fixed priors
      a[dept] ~ dnorm(0, 1),
      sigma_gender ~ dexp(1),
      Rho_gender ~ dlkjcorr(4)
    ),
    data = dat_list,
    chains = 4,
    cores = 4,
    iter = 5000,
    log_lik = T
  )
)

system.time(
  m14m3.noncentered <- ulam(
    alist(
      admit ~ dbinom(applications, p),
      logit(p) <- a[dept] + beta[gender, dept],
      # adaptive prior - non-centered
      transpars > matrix[gender, 6]:beta <-
        compose_noncentered(sigma_gender, L_Rho_gender, z_gender),
      matrix[6, gender]:z_gender ~ dnorm(0, 1),
      # fixed priors
      a[dept] ~ dnorm(0, 1),
      vector[6]:sigma_gender ~ dexp(1),
      cholesky_factor_corr[6]:L_Rho_gender ~ lkj_corr_cholesky(2),
      # compute ordinary correlation matrixes from Cholskey factors
      gq > matrix[6, 6]:Rho_gender <<-
        multiply_lower_tri_self_transpose(L_Rho_gender)
    ),
    data = dat_list,
    chains = 4,
    cores = 4,
    iter = 5000,
    log_lik = T
  )
)

compare(m14m3.centered,m14m3.noncentered)

neff_nc <- precis(m14m3.noncentered,pars=c("beta"), depth = 3)$n_eff
neff_c <- precis(m14m3.centered,pars=c("beta"), depth = 3)$n_eff
plot( neff_c , neff_nc , xlab="centered (default)" ,
ylab="non-centered (cholesky)" , lwd=1.5, xlim = c(0,5000), ylim = c(0,5000))
abline(a=0,b=1,lty=2)
```

Non-centered model was twice as fast. Way more effective samples in the non-centered model

# PDF
### 1. Revisit the Bangladesh fertility data, `data(bangladesh)`. Fit a model with both varying intercepts by `district_id` and varying slopes of `urban` (as a 0/1 indicator variable) by `district_id`. You are still predicting `use.contraception`. Inspect the correlation between the slopes and the intercepts. Can you interpret this correlation in terms of what it tells you about the pattern of contraceptive use in the sample? It might help to play the varying effect estimates for both the intercepts and slops, by district. Then you can visualize the correlation and maybe more easily think throught what it means to have a particular correlation. Plotting predicted proportion of women using contraception, in each district, with urban women on one axis and rural on the other might also help/

```{r}
data("bangladesh")
d <- bangladesh

dat_list <- list(
  district = as.integer(as.factor(d$district)),
  contraception = d$use.contraception,
  urban = d$urban
)

# Make it like the model m14.1
model1.1 <- ulam(
  alist(
    contraception ~ binomial(1,p),
    logit(p) <- alpha[district] + beta[district]*urban,
    c(alpha,beta)[district] ~ multi_normal(c(a,b), Rho, Sigma),
    a ~ dnorm(0,1),
    b ~ dnorm(0,1),
    Rho ~ lkj_corr(2),
    Sigma ~ dexp(1)
  ),
  data = dat_list,
  chains = 4,
  cores = 4,
  iter = 5000,
  log_lik = T
)

precis(model1.1, depth = 3, pars = c("a","b","Rho","Sigma"))

post <- extract.samples(model1.1)

dens( post$Rho[,1,2] )

a <- apply(post$alpha, 2, mean)
b <- apply(post$beta, 2, mean)
plot(
  a ,
  b ,
  xlab = "intercept" ,
  ylab = "slope" ,
  pch = 16 ,
  col = rangi2 ,
  ylim = c(min(b) - 0.1 , max(b) + 0.1) ,
  xlim = c(min(a) - 0.1 , max(a) + 0.1)
)

# compute posterior mean bivariate Gaussian
Mu_est <- c( mean(post$a) , mean(post$b) )
rho_est <- mean( post$Rho[,1,2] )
sa_est <- mean( post$Sigma[,1] )
sb_est <- mean( post$Sigma[,2] )
cov_ab <- sa_est*sb_est*rho_est
Sigma_est <- matrix( c(sa_est^2,cov_ab,cov_ab,sb_est^2) , ncol=2 )
# draw contours
library(ellipse)
for ( l in c(0.1,0.3,0.5,0.8,0.99) )
lines(ellipse(Sigma_est,centre=Mu_est,level=l),
col=col.alpha("black",0.2))
abline(h=0)
abline(v=0)

rural <- inv_logit(a)
urban <- inv_logit(a + b)

plot(
  urban ,
  rural ,
  xlab = "urban" ,
  ylab = "rural" ,
  pch = 16 ,
  col = rangi2 ,
  ylim = c(0,1) ,
  xlim = c(0,1)
)

abline(h=0.5)
abline(v=0.5)
```

Looks like being in an urban area increases probability of using contraception. Negative correlation between alpha and beta though. Rural use looks to be highly variable no matter what the urban use is.


### 2. Now consider the predictor variables `age.centered` and `living.children`, also contained in `data(bangladesh)`. Suppose that age influences contraceptive use (changing attitudes) and number of children (older people have had more time to have kids). Number of children may also directly influence contraceptive use. Draw a DAG that reflects these hypothetical relationships. Then build models needed to evaulate the DAG. You will need at least two models. Retain `district` and `urban` as in Problem 1. What do you conclude about the causal influence of age and children?

```{r}
library(dagitty)
dag2 <- dagitty( "dag {
K -> C
A -> C
A -> K
}")
coordinates(dag2) <- list( x=c(C=1,A=0,K=2) , y=c(C=0,A=1,K=1))
drawdag(dag2)
```

```{r}
dat_list <- list(
  district = as.integer(as.factor(d$district)),
  contraception = d$use.contraception,
  urban = d$urban,
  age = standardize(d$age.centered),
  kids = standardize(d$living.children)
)

m2.1 <- ulam(
  alist(
    contraception ~ binomial(1,p),
    logit(p) <- alpha[district] + betaD[district]*urban + betaA*age,
    c(alpha,betaD)[district] ~ multi_normal(c(a,b), Rho, Sigma),
    a ~ dnorm(0,1),
    b ~ dnorm(0,1),
    betaA ~ dnorm(0,1),
    Rho ~ lkj_corr(2),
    Sigma ~ dexp(1)
  ),
  data = dat_list,
  chains = 4,
  cores = 4,
  iter = 5000,
  log_lik = T
)

precis(m2.1)

m2.2 <- ulam(
  alist(
    contraception ~ binomial(1,p),
    logit(p) <- alpha[district] + betaD[district]*urban + betaK*kids,
    c(alpha,betaD)[district] ~ multi_normal(c(a,b), Rho, Sigma),
    a ~ dnorm(0,1),
    b ~ dnorm(0,1),
    betaK ~ dnorm(0,1),
    Rho ~ lkj_corr(2),
    Sigma ~ dexp(1)
  ),
  data = dat_list,
  chains = 4,
  cores = 4,
  iter = 5000,
  log_lik = T
)

precis(m2.2)

m2.3 <- ulam(
  alist(
    contraception ~ binomial(1,p),
    logit(p) <- alpha[district] + betaD[district]*urban + betaA*age + betaK*kids,
    c(alpha,betaD)[district] ~ multi_normal(c(a,b), Rho, Sigma),
    a ~ dnorm(0,1),
    b ~ dnorm(0,1),
    betaA ~ dnorm(0,1),
    betaK ~ dnorm(0,1),
    Rho ~ lkj_corr(2),
    Sigma ~ dexp(1)
  ),
  data = dat_list,
  chains = 4,
  cores = 4,
  iter = 5000,
  log_lik = T
)

precis(m2.3)

```

Kids seem to be the deciding factor in contraception use. Having kids makes you not want to have kids
