---
title: "Germination Modeling"
author: "John D."
date: "10_7_2020"
output:
  github_document:
    pandoc_args: --webtex
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=F}
library(rethinking)
library(tidyverse)
```       

```{r}
dat <- read_csv("light_round1_tall.csv")
head(dat)
```
# Filter down data
```{r}
## Explode data set, 1 row per seed

dat.explode <- data.frame("population" = NA,
                          "temperature" = NA,
                          "water_potential" = NA,
                          "germ" = NA,
                          "days_to_event" = NA)
dat.explode <- dat.explode[-1,]
### Germinated seeds
dat.small <- dat %>%
  filter(germ != 0) %>%
  select("population" = pops, "temperature" = temps, 
         "water_potential" = wps, germ, "days_to_event" = day)

for(i in 1:nrow(dat.small)){
  dat.explode <- rbind(dat.explode,
                       dat.small %>% slice(rep(i:i, each = dat.small[[i,4]]))
  )
}

dat.explode$germ <- 1

### Ungerminated seeds
duds_per_trt <- dat %>%
  group_by(pops, temps, wps) %>%
  slice(n()) %>%
  ungroup() %>%
  mutate(duds = total_seeds -cumulative_germ,
         duds = ifelse(duds <0, 0, duds),
         germ = 0) %>%
  select("population" = pops, "temperature" = temps, 
         "water_potential" = wps, germ,
         "days_to_event" = day, duds)

for(i in 1:nrow(duds_per_trt)){
  dat.explode <- rbind(dat.explode,
                       duds_per_trt[,-6] %>% slice(rep(i:i, each = duds_per_trt[[i,6]]))
  )
}

dim(dat.explode)
head(dat.explode)
```

```{r}
## Make data list
dat_list <- list(
  Germ = as.integer(dat.explode$germ),
  Day = dat.explode$days_to_event,
  Water = dat.explode$water_potential,
  Temp = dat.explode$temperature,
  Pop = as.integer(as.factor(dat.explode$population))
)
```

```{r}
## Modeling
### Basic, will start using the cat example
model_1 <- ulam(
  alist(
    Day | Germ == 1 ~ exponential(lambda),
    Day | Germ == 0 ~ custom(exponential_lccdf(!Y | lambda)),
    lambda <- 1.0 / mu,
    log(mu) <- alpha_pop[Pop],
    alpha_pop[Pop] ~ normal(0, 1)
  ),
  data = dat_list ,
  chains = 4 ,
  cores = 4
)

precis(model_1, depth = 2)
post <- extract.samples(model_1)
post$D <- exp(post$alpha_pop)
precis( post , 2 )
```
This modeling using a Poisson distribution appears to be showing the average days per a population to germinate. They are extremely high, most likely due to zero inflation. Time to convert to Zero inflated Poisson.

```{r, eval=F}
### Converting cat example to zero inflated Poisson
model_2 <- ulam(
  alist(
    Day | Germ == 1 ~  custom( log1m(p) + poisson_lpmf(Germ|lambda) ),
    Day | Germ == 0 ~ custom(log_mix( p , 0 , poisson_lpmf(0|lambda) ) ),
    logit(p) <- ap[Pop],
    log(lambda) <- al[Pop],
    ap[Pop] ~ dnorm(0,1),
    al[Pop] ~ dnorm(1,3)
  ),
  data = dat_list ,
  chains = 4 ,
  cores = 4
)
### This would take ~16 hours to run on full set. will try on just 2 pops
```

```{r}
dat2 <- dat.explode %>%
  filter(population == "STDI" | population == "CAAM")

dat_list2 <- list(
  Germ = as.integer(dat2$germ),
  Day = dat2$days_to_event,
  Water = dat2$water_potential,
  Temp = dat2$temperature,
  Pop = as.integer(as.factor(dat2$population))
)

model_2 <- ulam(
  alist(
    Day | Germ == 1 ~  custom( log1m(p) + poisson_lpmf(Germ|lambda) ),
    Day | Germ == 0 ~ custom(log_mix( p , 0 , poisson_lpmf(0|lambda) ) ),
    logit(p) <- ap[Pop],
    log(lambda) <- al[Pop],
    ap[Pop] ~ dnorm(0,1),
    al[Pop] ~ dnorm(1,3)
  ),
  data = dat_list2 ,
  chains = 4 ,
  cores = 4
)

precis(model_2, depth = 2)
post <- extract.samples(model_2)
mean( inv_logit( post$ap ) ) # probability dead
mean( exp( post$al ) ) # approximate seeds per a day
```

Seems like it's kinda working, not exactly a germination curve though. Try treating like an ordered categorical outcome where probability of germinating increases with each day. Need to also fit in zero inflation

```{r}
simplehist( dat.explode$days_to_event , xlim=c(1,28) , xlab="day of germination" )
```

```{r}
### Look at STDI
# discrete proportion of each response value
pr_k <- table( dat2 %>% filter(population == "STDI" ) %>% select(days_to_event)) / nrow(dat2 %>% filter(population == "STDI" ))
# cumsum converts to cumulative proportions
cum_pr_k <- cumsum( pr_k )
# plot
plot( c(1,2,3,4,5,6,7,9,11,13,17,19,21,24,28) , cum_pr_k , type="b" , xlab="germination" ,
ylab="cumulative proportion" , ylim=c(0,1) )
```

```{r}
# Can only use seeds that germinated
dat3 <- dat2 %>% filter(germ != 0)
dat_list3 <- list(
  Germ = as.integer(dat3$germ),
  Day = dat3$days_to_event,
  Water = dat3$water_potential,
  Temp = dat3$temperature,
  Pop = as.integer(as.factor(dat3$population))
)

model_3 <- ulam(
  alist(
    Day ~ dordlogit(phi , cutpoints),
    phi <- alpha_pop[Pop] + beta_water[Pop]*Water + beta_temp[Pop]*Temp,
    alpha_pop[Pop] ~ dnorm(0 , 0.5),
    beta_water[Pop] ~ dnorm(0 , 0.5),
    beta_temp[Pop] ~ dnorm(0 , 0.5),
    cutpoints ~ dnorm(0 , 1.5)
  ) ,
  data = dat_list3 ,
  chains = 4 ,
  cores = 4
)
precis(model_3, depth = 2)
round( inv_logit(coef(model_3)) , 3 )
```

