---
title: "Ch.5 HW"
output:
  html_document:
    keep_md: yes
editor_options:
  chunk_output_type: inline
---

## Easy

### 5E1. Which of the linear models below are multiple linear regressions?
  (1) µi = α + βxi
  (2) µi = βxxi + βzzi
  (3) µi = α + β(xi − zi)
  (4) µi = α + βxxi + βzz
  
2 and 4

### 5E3. Write down a multiple regression to evaluate the claim: Neither amount of funding nor size of laboratory is by itself a good predictor of time to PhD degree; but together these variables are both positively associated with time to degree. Write down the model definition and indicate which side of zero each slope parameter should be on.

$$
time2phd_i \sim (\mu_i, \sigma) \\
mu_i = \alpha + \beta Funding*Funding_i + \beta Size*Size_i \\
\alpha \sim Norm(6,2) \\
\beta Funding \sim Norm(0,0.5) \\
\beta Size \sim Norm(0,0.5 \\
\sigma \sim exp(1) \\
$$

## Medium

### 5M2. Invent your own example of a masked relationship. An outcome variable should be correlated with both predictor variables, but in opposite directions. And the two predictor variables should be correlated with one another.

Win rate at a video game is influenced by amount of time played and time of day

More time played leads to winning more

During evening better players come online making winning harder

If you play more you will probaby also play at later times of the day

### 5M3. It is sometimes observed that the best predictor of fire risk is the presence of firefighters—States and localities with many firefighters also have more fires. Presumably firefighters do not cause fires. Nevertheless, this is not a spurious correlation. Instead fires cause firefighters. Consider the same reversal of causal inference in the context of the divorce and marriage data. How might a high divorce rate cause a higher marriage rate? Can you think of a way to evaluate this relationship, using multiple regression?

More people getting divorced leads to more single people which leads to more people available to get married.

$$
MariageRate \sim DivorceRate + AgeAtMarriage
$$

### 5M4. In the divorce data, States with high numbers of Mormons (members of The Church of Jesus Christ of Latter-day Saints, LDS) have much lower divorce rates than the regression models expected. Find a list of LDS population by State and use those numbers as a predictor variable, predicting divorce rate using marriage rate, median age at marriage, and percent LDS population (possibly standardized). You may want to consider transformations of the raw percent LDS variable.

```{r}
library(XML)
library(RCurl)
library(rlist)
library(tidyverse)
library(rethinking)
data("WaffleDivorce")
d <- WaffleDivorce
theurl <- getURL("https://www.worldatlas.com/articles/mormon-population-by-state.html",.opts = list(ssl.verifypeer = FALSE) )
tables <- readHTMLTable(theurl)
tables <- list.clean(tables, fun = is.null, recursive = FALSE)
n.rows <- unlist(lapply(tables, function(t) dim(t)[1]))
mormons <- tables[[which.max(n.rows)]]
colnames(mormons) <- c("Rank","State","Trash","Trash2","Pct")
str(mormons)
mormons <- mormons[,c(2,5)]
mormons$State <- as.character(mormons$State)
mormons$Pct <- as.numeric(sub("%","",as.character(mormons$Pct)))
str(mormons)
d <- left_join(d,mormons, by = c("Location" = "State"))

dstd <- d %>%
  mutate(MedianAgeMarriage = scale(MedianAgeMarriage),
         Pct = scale(log(Pct)),
         Marriage = scale(Marriage),
         Divorce = scale(Divorce))
head(dstd)
```

```{r}
fit <- quap(
  alist(
    Divorce ~ dnorm(mu, sigma),
    mu <- a + bA*MedianAgeMarriage + bM*Marriage +bP*Pct,
    a ~ dnorm(0,0.5),
    bA ~ dnorm(0,0.5),
    bM ~ dnorm(0,0.5),
    bP ~ dnorm(0,0.5),
    sigma ~ dexp(1)),
  data = dstd)
precis(fit)
```

```{r}
# call link without specifying new data
# so it uses original data
mu <- link( fit )
# summarize samples across cases
mu_mean <- apply( mu , 2 , mean )
mu_PI <- apply( mu , 2 , PI )
# simulate observations
# again no new data, so uses original data
D_sim <- sim( fit , n=1e4 )
D_PI <- apply( D_sim , 2 , PI )

plot( mu_mean ~ dstd$Divorce , col=rangi2 , ylim=range(mu_PI) ,
xlab="Observed divorce" , ylab="Predicted divorce" )
abline( a=0 , b=1 , lty=2 )
for ( i in 1:nrow(d) ) lines( rep(dstd$Divorce[i],2) , mu_PI[,i] , col=rangi2 )
```


## Hard
```{r}
data(foxes)
d <- foxes
str(d)
summary(d)
```

### 5H1. Fit two bivariate Gaussian regressions, using quap: (1) body weight as a linear function of territory size (area), and (2) body weight as a linear function of groupsize. Plot the results of these regressions, displaying the MAP regression line and the 95% interval of the mean. Is either variable important for predicting fox body weight?

```{r}
H511 <- quap(
  alist(
    weight ~ dnorm(mu, sigma),
    mu <- a + bA*area,
    a ~ dnorm(3,2),
    bA ~ dnorm(0,1),
    sigma ~ dexp(1)),
  data = d
)
precis(H511)
```
```{r}
xseq <- seq( from=min(d$area)-0.15 , to=max(d$area)+0.15 , length.out=100 )
mu <- link( H511 , data=list(area=xseq) )
mu_mean <- apply(mu,2,mean)
mu_PI <- apply(mu,2,HPDI, .95)
plot( weight ~ area , data=d )
lines( xseq , mu_mean , lwd=2 )
shade( mu_PI , xseq )
```
 

```{r}
H512 <- quap(
  alist(
    weight ~ dnorm(mu, sigma),
    mu <- a + bS*groupsize,
    a ~ dnorm(4,3),
    bS ~ dnorm(0,1),
    sigma ~ dexp(1)),
  data = d
)
precis(H512)
```

```{r}
xseq <- seq( from=min(d$groupsize)-0.15 , to=max(d$groupsize)+0.15 , length.out=100 )
mu <- link( H512 , data=list(groupsize=xseq) )
mu_mean <- apply(mu,2,mean)
mu_PI <- apply(mu,2,HPDI, .95)
plot( weight ~ groupsize , data=d )
lines( xseq , mu_mean , lwd=2 )
shade( mu_PI , xseq )
```

Maybe groupsize

### 5H2. Now fit a multiple linear regression with weight as the outcome and both area and groupsize as predictor variables. Plot the predictions of the model for each predictor, holding the other predictor constant at its mean. What does this model say about the importance of each variable? Why do you get different results than you got in the exercise just above?

```{r}
H521 <- quap(
  alist(
    weight ~ dnorm(mu, sigma),
    mu <- a + bA*area + bS*groupsize,
    a ~ dnorm(3,2),
    bA ~ dnorm(0,1),
    bS ~ dnorm(0,1),
    sigma ~ dexp(1)),
  data = d
)
precis(H521)
```

```{r}
xseq <- seq( from=min(d$area)-0.15 , to=max(d$area)+0.15 , length.out=100 )
mu <- link( H521 , data=data.frame( area=xseq , groupsize=0 ) )
mu_mean <- apply(mu,2,mean)
mu_PI <- apply(mu,2,PI)
plot( NULL , xlim=range(d$area) , ylim=range(d$weight) )
lines( xseq , mu_mean , lwd=2 )
shade( mu_PI , xseq )
```
```{r}
xseq <- seq( from=min(d$groupsize)-0.15 , to=max(d$groupsize)+0.15 , length.out=100 )
mu <- link( H521 , data=data.frame( area=0 , groupsize=xseq ) )
mu_mean <- apply(mu,2,mean)
mu_PI <- apply(mu,2,PI)
plot( NULL , xlim=range(d$groupsize) , ylim=range(d$weight) )
lines( xseq , mu_mean , lwd=2 )
shade( mu_PI , xseq )
```

Opposite directions

```{r}
pairs(d)
```

