---
title: "01_24_2020_HW"
author: "John D."
date: "1/23/2020"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=F}
library(tidyverse)
library(rethinking)
```

## Medium
### 12M3. Re-estimate the basic Reed frog varying intercept model, but now using a Cauchy distribution in place of the Gaussian distribution for the varying intercepts. That is, fit this model:
$$
   s_i \sim Binomial(n_i, p_i) \\
   logit(p_i) = \alpha_{TANK[i]} \\
   \alpha_{TANK} \sim Cauchy(\alpha, \sigma) \\
   \alpha \sim Normal(0, 1) \\
   \sigma \sim HalfCauchy(0, 1)
$$

### Compare the posterior means of the intercepts, αtank, to the posterior means produced in the chapter, using the customary Gaussian prior. Can you explain the pattern of differences?

```{r}
data(reedfrogs)
d <- reedfrogs
str(d)

# make the tank cluster variable
d$tank <- 1:nrow(d)

dat <- list(S = d$surv,
            N = d$density,
            tank = d$tank)

# Original model
m_normal <- ulam(
  alist(
    S ~ dbinom(N , p) ,
    logit(p) <- a[tank] ,
    a[tank] ~ dnorm(a_bar , sigma) ,
    a_bar ~ dnorm(0 , 1.5) ,
    sigma ~ dexp(1)
  ),
  data = dat ,
  chains = 4 ,
  log_lik = TRUE,
  cores = 4
)

# Cauchy model
m_cauchy <- ulam(
  alist(
    S ~ dbinom(N , p) ,
    logit(p) <- a[tank] ,
    a[tank] ~ dcauchy(a_bar , sigma) ,
    a_bar ~ dnorm(0 , 1.5) ,
    sigma ~ dcauchy(0, 1)
  ),
  data = dat ,
  chains = 4 ,
  log_lik = TRUE,
  cores = 4
)

compare(m_normal,m_cauchy)

precis(m_normal)
precis(m_cauchy)

post_normal <- extract.samples(m_normal)
post_cauchy <- extract.samples(m_cauchy)

max(post_normal$a)
max(post_cauchy$a)

dens(post_normal$a)
dens(post_cauchy$a)

dens(post_cauchy$a)
dens(post_normal$a, add = T, col = 'blue')
```

The cauchy distribution has a lot longer tails than the normal distribution making the range of estimates larger.

### 12M4. Fit the following cross-classified multilevel model to the chimpanzees data:
$$
L_i \sim Binomial(1, p_i) \\
logit(p_i) = \alpha_{actor[i]} + \alpha_{block[i]} + (\beta_P + \beta_{PC}C_i)P_i \\
\alpha_{actor} \sim Normal(\alpha, \sigma_{actor}) \\
\alpha_{block} \sim Normal(\gamma, \sigma_{block}) \\
\alpha, \gamma, \beta_P, \beta_{PC} \sim Normal(0, 10) \\
\sigma_{actor}, \sigma_{block} \sim HalfCauchy(0, 1)
$$
```{r}
data(chimpanzees)
d <- chimpanzees
d$treatment <- 1 + d$prosoc_left + 2 * d$condition
dat_list <- list(
  pulled_left = d$pulled_left,
  actor = d$actor,
  block_id = d$block,
  treatment = as.integer(d$treatment)
)

m_chapter <- ulam(
  alist(
    pulled_left ~ dbinom(1 , p) ,
    logit(p) <- a[actor] + g[block_id] + b[treatment],
    b[treatment] ~ dnorm(0 , 0.5),
    # adaptive priors
    a[actor] ~ dnorm(a_bar , sigma_a),
    g[block_id] ~ dnorm(0 , sigma_g),
    # hyper-priors
    a_bar ~ dnorm(0 , 1.5),
    sigma_a ~ dexp(1),
    sigma_g ~ dexp(1)
  ) ,
  data = dat_list ,
  chains = 4 ,
  cores = 4 ,
  log_lik = TRUE
)

dat_list2 <- list(
  pulled_left = d$pulled_left,
  actor = d$actor,
  block_id = d$block,
  prosocial = as.integer(d$prosoc_left),
  condition = as.integer(d$condition)
) 
m_problem <- ulam(
  alist(
    pulled_left ~ dbinom(1 , p) ,
    logit(p) <- a[actor] + g[block_id] + (bP + bPC * condition) * prosocial,
    # adaptive priors
    a[actor] ~ dnorm(a_bar , sigma_a),
    g[block_id] ~ dnorm(g_bar , sigma_g),
    # hyper-priors
    c(a_bar,g_bar,bP,bPC) ~ dnorm(0 , 10),
    c(sigma_a,sigma_g) ~ dcauchy(0,1)
  ) ,
  data = dat_list2 ,
  chains = 4 ,
  cores = 4 ,
  log_lik = TRUE
)

coeftab(m_chapter, m_problem)

plot(m_chapter)
plot(m_problem)

post_c <- extract.samples(m_chapter)
post_p <- extract.samples(m_problem)
prob_c <- data.frame(inv_logit(post_c$a))
prob_p <- data.frame(inv_logit(post_p$a))
par(mfrow = c(1,2))
plot(precis(prob_c))
plot(precis(prob_p))

neff_c <- data.frame(neff_c = precis( m_chapter , depth=2 )[['n_eff']])
neff_c$param <- rownames(precis( m_chapter , depth=2 ))
neff_p <- data.frame(neff_p = precis( m_problem , depth=2 )[['n_eff']])
neff_p$param <- rownames(precis( m_problem , depth=2 ))

par_names <- unique(c(neff_c$param,neff_p$param))
neff_table <- data.frame(param = par_names)
neff_table <- left_join(neff_table,neff_c) %>%
  left_join(neff_p)
neff_table <- as.matrix(neff_table[,-1])
rownames(neff_table) <- par_names
round(t(neff_table))
```

### Each of the parameters in those comma-separated lists gets the same independent prior. Compare the posterior distribution to that produced by the similar cross-classified model from the chapter. Also compare the number of effective samples. Can you explain the differences?

The formula from the homework has a lot less effective samples compared to the formula from the chapter. Also a lot of uncertainty in a_bar and g_bar leads the estimate for each chimp to be abysmal.

## Hard
### 12H2. Return to the Trolley data, `data(Trolley)`, from Chapter 12. Define and fit a varying intercepts model for these data. Cluster intercepts on individual participants, as indicated by the unique values in the `id` variable. Include `action`, `intention`, and `contact` as ordinary terms. Compare the varying intercepts model and a model that ignores individuals, using both WAIC and posterior predictions. What is the impact of individual variation in these data?

```{r}
data(Trolley)
d <- Trolley
head(Trolley)
dat <- list(
  R = d$response,
  A = d$action,
  I = d$intention,
  C = d$contact,
  ID = as.numeric(d$id)
)

length(unique(dat$ID)) # 331 people

m_old <- ulam(
  alist(
    R ~ dordlogit(phi , cutpoints),
    phi <- bA * A + bC * C + BI * I ,
    BI <- bI + bIA * A + bIC * C ,
    c(bA, bI, bC, bIA, bIC) ~ dnorm(0 , 0.5),
    cutpoints ~ dnorm(0 , 1.5)
  ) ,
  data = dat ,
  chains = 4 ,
  cores = 4,
  log_lik = T
)

m_new <- ulam(
  alist(
    R ~ dordlogit(phi , cutpoints),
    phi <- a[ID] + bA * A + bC * C + BI * I ,
    BI <- bI + bIA * A + bIC * C ,
    c(bA, bI, bC, bIA, bIC) ~ dnorm(0 , 0.5),
    a[ID] ~ dnorm(a_bar,sigma_a),
    a_bar ~ dnorm(0,1.5),
    sigma_a ~ dexp(1),
    cutpoints ~ dnorm(0 , 1.5)
  ) ,
  data = dat ,
  chains = 4 ,
  cores = 4,
  log_lik = T
)

precis(m_old)
precis(m_new)
compare(m_old,m_new)
coeftab(m_old,m_new)
par(mfrow = c(1,1))
plot(coeftab(m_old,m_new), pars = c("bIC","bIA","bC","bI","bA"))
```

New model has a way lower WAIC


### 12H3. The Trolley data are also clustered by `story`, which indicates a unique narrative for each vignette. Define and fit a cross-classified varying intercepts model with both `id` and `story`. Use the same ordinary terms as in the previous problem. Compare this model to the previous models. What do you infer about the impact of different stories on responses?

```{r}
dat <- list(
  R = d$response,
  A = d$action,
  I = d$intention,
  C = d$contact,
  ID = as.numeric(d$id),
  S = as.numeric(d$story)
)

m_newest <- ulam(
  alist(
    R ~ dordlogit(phi , cutpoints),
    phi <- a[ID] + s[S] + bA * A + bC * C + BI * I ,
    BI <- bI + bIA * A + bIC * C ,
    c(bA, bI, bC, bIA, bIC) ~ dnorm(0 , 0.5),
    a[ID] ~ dnorm(a_bar,sigma_a),
    s[S] ~ dnorm(0, sigma_s),
    a_bar ~ dnorm(0,1.5),
    c(sigma_a, sigma_s) ~ dexp(1),
    cutpoints ~ dnorm(0 , 1.5)
  ) ,
  data = dat ,
  chains = 4 ,
  cores = 4,
  log_lik = T
)

precis(m_newest)
compare(m_old,m_new,m_newest)
coeftab(m_old,m_new,m_newest)
par(mfrow = c(1,1))
plot(coeftab(m_old,m_new,m_newest), pars = c("bIC","bIA","bC","bI","bA"))
```

The newest model appears to be the best, but it's estimates seem worse.