---
title: "01_24_2020_Notes"
author: "John D."
date: "1/24/2020"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(rethinking)
```

# 13 Models With Memory
## 13.1. Example: Multilevel tadpoles

```{r}
data(reedfrogs)
d <- reedfrogs
str(d)
```

```{r}
# make the tank cluster variable
d$tank <- 1:nrow(d)

dat <- list(S = d$surv,
            N = d$density,
            tank = d$tank)

# approximate posterior
m13.1 <- ulam(
  alist(S ~ dbinom(N , p) ,
        logit(p) <- a[tank] ,
        a[tank] ~ dnorm(0 , 1.5)),
  data = dat ,
  chains = 1 ,
  log_lik = TRUE
)
```

```{r}
precis(m13.1, depth = 2)
```

```{r}
m13.2 <- ulam(
  alist(
    S ~ dbinom(N , p) ,
    logit(p) <- a[tank] ,
    a[tank] ~ dnorm(a_bar , sigma) ,
    a_bar ~ dnorm(0 , 1.5) ,
    sigma ~ dexp(1)
  ),
  data = dat ,
  chains = 1 ,
  log_lik = TRUE
)
```

```{r}
precis(m13.2, depth = 2)

compare( m13.1 , m13.2 )
```

```{r}
# extract Stan samples
post <- extract.samples(m13.2)

# compute median intercept for each tank
# also transform to probability with logistic
d$propsurv.est <- logistic(apply(post$a , 2 , mean))

# display raw proportions surviving in each tank
plot(
  d$propsurv ,
  ylim = c(0, 1) ,
  pch = 16 ,
  xaxt = "n" ,
  xlab = "tank" ,
  ylab = "proportion survival" ,
  col = rangi2
)
axis( 1 , at=c(1,16,32,48) , labels=c(1,16,32,48) )

# overlay posterior means
points( d$propsurv.est )

# mark posterior mean probability across tanks
abline( h=mean(inv_logit(post$a_bar)) , lty=2 )

# draw vertical dividers between tank densities
abline( v=16.5 , lwd=0.5 )
abline( v=32.5 , lwd=0.5 )
text( 8 , 0 , "small tanks" )
text( 16+8 , 0 , "medium tanks" )
text( 32+8 , 0 , "large tanks" )
```

```{r}
# show first 100 populations in the posterior
plot(
  NULL ,
  xlim = c(-3, 4) ,
  ylim = c(0, 0.35) ,
  xlab = "log-odds survive" ,
  ylab = "Density"
)

for (i in 1:100)
  curve(dnorm(x, post$a_bar[i], post$sigma[i]) ,
        add = TRUE ,
        col = col.alpha("black", 0.2))

# sample 8000 imaginary tanks from the posterior distribution
sim_tanks <- rnorm(8000 , post$a_bar , post$sigma)

# transform to probability and visualize
dens(inv_logit(sim_tanks) , lwd = 2 , adj = 0.1)
```

## 13.2. Varying effects and the underfitting/overfitting trade-off
### 13.2.1. The model
### 13.2.2. Assign values to the parameters

```{r}
a_bar <- 1.5
sigma <- 1.5
nponds <- 60
Ni <- as.integer( rep( c(5,10,25,35) , each=15 ) )
```

```{r}
set.seed(5005)
a_pond <- rnorm( nponds , mean=a_bar , sd=sigma )
a_pond
```

```{r}
dsim <- data.frame( pond=1:nponds , Ni=Ni , true_a=a_pond )
head(dsim)
```

###  13.2.3. Simulate survivors

```{r}
dsim$Si <- rbinom( nponds , prob=logistic(dsim$true_a) , size=dsim$Ni )
head(dsim)
```

### 13.2.4. Compute the no-pooling estimates

```{r}
dsim$p_nopool <- dsim$Si / dsim$Ni
head(dsim)
```

### 13.2.5. Compute the partial-pooling estimates

```{r}
dat <- list(Si = dsim$Si ,
            Ni = dsim$Ni ,
            pond = dsim$pond)

m13.3 <- ulam(
  alist(
    Si ~ dbinom(Ni , p),
    logit(p) <- a_pond[pond],
    a_pond[pond] ~ dnorm(a_bar , sigma),
    a_bar ~ dnorm(0 , 1.5),
    sigma ~ dexp(1)
  ),
  data = dat ,
  chains = 4,
  cores = 3
)

precis(m13.3 , depth = 2)
```

```{r}
post <- extract.samples( m13.3 )
dsim$p_partpool <- apply( inv_logit(post$a_pond) , 2 , mean )
```

```{r}
dsim$p_true <- inv_logit( dsim$true_a )
```

```{r}
nopool_error <- abs( dsim$p_nopool - dsim$p_true )
partpool_error <- abs( dsim$p_partpool - dsim$p_true )
nopool_avg <- aggregate(nopool_error,list(dsim$Ni),mean)
partpool_avg <- aggregate(partpool_error,list(dsim$Ni),mean)
```

```{r}
head(dsim)
```

```{r}
plot(
  1:60 ,
  nopool_error ,
  xlab = "pond" ,
  ylab = "absolute error" ,
  col = rangi2 ,
  pch = 16
)
points(1:60 , partpool_error)
```

### Overthinking: Repeating the pond simulation

```{r}
a <- 1.4
sigma <- 1.5
nponds <- 60
Ni <- as.integer(rep(c(5, 10, 25, 35) , each = 15))
a_pond <- rnorm(nponds , mean = a , sd = sigma)
dsim <- data.frame(pond = 1:nponds ,
                   Ni = Ni ,
                   true_a = a_pond)
dsim$Si <-
  rbinom(nponds, prob = inv_logit(dsim$true_a), size = dsim$Ni)
dsim$p_nopool <- dsim$Si / dsim$Ni
newdat <- list(Si = dsim$Si,
               Ni = dsim$Ni,
               pond = 1:nponds)
m13.3new <- stan(fit = m13.3@stanfit ,
                 data = newdat ,
                 chains = 1)
post <- extract.samples(m13.3new)
dsim$p_partpool <- apply(inv_logit(post$a_pond) , 2 , mean)
dsim$p_true <- inv_logit(dsim$true_a)
nopool_error <- abs(dsim$p_nopool - dsim$p_true)
partpool_error <- abs(dsim$p_partpool - dsim$p_true)
plot(
  1:60 ,
  nopool_error ,
  xlab = "pond" ,
  ylab = "absolute error" ,
  col = rangi2 ,
  pch = 16
)
points(1:60 , partpool_error)
```

## 13.3. More than one type of cluster
### 13.3.1. Multilevel chimpanzees

```{r}
data(chimpanzees)
d <- chimpanzees
d$treatment <- 1 + d$prosoc_left + 2 * d$condition
dat_list <- list(
  pulled_left = d$pulled_left,
  actor = d$actor,
  block_id = d$block,
  treatment = as.integer(d$treatment)
)
set.seed(13)
m13.4 <- ulam(
  alist(
    pulled_left ~ dbinom(1 , p) ,
    logit(p) <- a[actor] + g[block_id] + b[treatment],
    b[treatment] ~ dnorm(0 , 0.5),
    # adaptive priors
    a[actor] ~ dnorm(a_bar , sigma_a),
    g[block_id] ~ dnorm(0 , sigma_g),
    # hyper-priors
    a_bar ~ dnorm(0 , 1.5),
    sigma_a ~ dexp(1),
    sigma_g ~ dexp(1)
  ) ,
  data = dat_list ,
  chains = 4 ,
  cores = 4 ,
  log_lik = TRUE
)
```

```{r}
precis( m13.4 , depth=2 )
plot( precis(m13.4,depth=2) ) # also plot
```

```{r}
set.seed(14)
m13.5 <- ulam(
  alist(
    pulled_left ~ dbinom(1 , p) ,
    logit(p) <- a[actor] + b[treatment] ,
    b[treatment] ~ dnorm(0 , 0.5),
    a[actor] ~ dnorm(a_bar , sigma_a),
    a_bar ~ dnorm(0 , 1.5),
    sigma_a ~ dexp(1)
  ) ,
  data = dat_list ,
  chains = 4 ,
  cores = 4 ,
  log_lik = TRUE
)

compare(m13.4, m13.5)
```

### 13.3.2. Even more clusters

```{r}
set.seed(15)
m13.6 <- ulam(
  alist(
    pulled_left ~ dbinom(1 , p) ,
    logit(p) <- a[actor] + g[block_id] + b[treatment] ,
    b[treatment] ~ dnorm(0 , sigma_b),
    a[actor] ~ dnorm(a_bar , sigma_a),
    g[block_id] ~ dnorm(0 , sigma_g),
    a_bar ~ dnorm(0 , 1.5),
    sigma_a ~ dexp(1),
    sigma_g ~ dexp(1),
    sigma_b ~ dexp(1)
  ) ,
  data = dat_list ,
  chains = 4 ,
  cores = 4 ,
  log_lik = TRUE
)

coeftab(m13.4, m13.6)

```

## 13.4. Divergent transitions and non-centered priors
### 13.4.1. The Devil’s Funnel

```{r}
m13x <- ulam(
  alist(v ~ normal(0, 3),
        x ~ normal(0, exp(v))),
  data = list(N = 1),
  chains = 4)
precis(m13x)
traceplot(m13x)
```

```{r}
m13y <- ulam(
  alist(v ~ normal(0, 3),
        z ~ normal(0, 1),
        gq> real[1]:x <<- z * exp(v)),
  data = list(N = 1) ,
  chains = 4
)
precis(m13y)
traceplot(m13y)
```

### 13.4.2. Non-centered chimpanzees

```{r}
set.seed(13)
m13.4b <- ulam( m13.4 , chains=4 , cores=4 , control=list(adapt_delta=0.99) )
divergent(m13.4b)
precis(m13.4b, depth = 2)
```

```{r}
set.seed(13)
m13.4nc <- ulam(
  alist(
    pulled_left ~ dbinom(1 , p) ,
    logit(p) <- a_bar + z[actor] * sigma_a + # actor intercepts
      x[block_id] * sigma_g + # block intercepts
      b[treatment] ,
    b[treatment] ~ dnorm(0 , 0.5),
    z[actor] ~ dnorm(0 , 1),
    x[block_id] ~ dnorm(0 , 1),
    a_bar ~ dnorm(0 , 1.5),
    sigma_a ~ dexp(1),
    sigma_g ~ dexp(1)
  ) ,
  data = dat_list ,
  chains = 4 ,
  cores = 4
)
 precis(m13.4nc, depth = 2)
```

```{r}
neff_c <- precis( m13.4 , depth=2 )[['n_eff']]
neff_nc <- precis( m13.4nc , depth=2 )[['n_eff']]
par_names <- rownames( precis( m13.4 , depth=2 ) )

neff_table <- cbind( neff_c , neff_nc )
rownames(neff_table) <- par_names
round(t(neff_table))
```

## 13.5. Multilevel posterior predictions
### 13.5.1. Posterior prediction for same clusters

```{r}
chimp <- 2
d_pred <- list(
  actor = rep(chimp, 4),
  treatment = 1:4,
  block_id = rep(1, 4)
)
p <- link(m13.4 , data = d_pred)
p_mu <- apply(p , 2 , mean)
p_ci <- apply(p , 2 , PI)

p_mu
p_ci

plot(
  NULL ,
  xlab = "treatment" ,
  ylab = "proportion pulled left" ,
  ylim = c(0, 1) ,
  xaxt = "n" ,
  xlim = c(1, 4)
)
axis(1 ,
     at = 1:4 ,
     labels = c("R/N", "L/N", "R/P", "L/P"))
lines(1:4 , p_mu)
shade(p_ci , 1:4)
```

```{r}
post <- extract.samples(m13.4)
str(post)
```

```{r}
dens( post$a[,5] )
dens( post$a[,2] )
dens( post$a[,7] )
par(mfrow = c(1,1))
```

```{r}
p_link <- function(treatment ,
                   actor = 1 ,
                   block_id = 1) {
  logodds <- with(post ,
                  a[, actor] + g[, block_id] + b[, treatment])
  return(inv_logit(logodds))
}
```

```{r}
p_raw <- sapply( 1:4 , function(i) p_link( i , actor=1 , block_id=1 ) )
p_mu <- apply( p_raw , 2 , mean )
p_ci <- apply( p_raw , 2 , PI )

p_mu
p_ci

plot(
  NULL ,
  xlab = "treatment" ,
  ylab = "proportion pulled left" ,
  ylim = c(0, 1) ,
  xaxt = "n" ,
  xlim = c(1, 4)
)
axis(1 ,
     at = 1:4 ,
     labels = c("R/N", "L/N", "R/P", "L/P"))
lines(1:4 , p_mu)
shade(p_ci , 1:4)
```

### 13.5.2. Posterior prediction for new clusters

```{r}
p_link_abar <- function(treatment) {
  logodds <- with(post , a_bar + b[, treatment])
  return(inv_logit(logodds))
}
```

```{r}
p_raw <- sapply(1:4 , function(i) p_link_abar(i))

p_mu <- apply(p_raw , 2 , mean)
p_ci <- apply(p_raw , 2 , PI)

plot(
  NULL ,
  xlab = "treatment" ,
  ylab = "proportion pulled left" ,
  ylim = c(0, 1) ,
  xaxt = "n" ,
  xlim = c(1, 4)
)
axis(1 ,
     at = 1:4 ,
     labels = c("R/N", "L/N", "R/P", "L/P"))
lines(1:4 , p_mu)
shade(p_ci , 1:4)
```

```{r}
a_sim <- with(post , rnorm(length(post$a_bar) , a_bar , sigma_a))
p_link_asim <- function(treatment) {
  logodds <- with(post , a_sim + b[, treatment])
  return(inv_logit(logodds))
}
p_raw_asim <- sapply(1:4 , function(i) p_link_asim(i))

p_mu_asim <- apply(p_raw_asim , 2 , mean)
p_ci_asim <- apply(p_raw_asim , 2 , PI)

plot(
  NULL ,
  xlab = "treatment" ,
  ylab = "proportion pulled left" ,
  ylim = c(0, 1) ,
  xaxt = "n" ,
  xlim = c(1, 4)
)
axis(1 ,
     at = 1:4 ,
     labels = c("R/N", "L/N", "R/P", "L/P"))
lines(1:4 , p_mu_asim)
shade(p_ci_asim , 1:4)
```

```{r}
plot(
  NULL ,
  xlab = "treatment" ,
  ylab = "proportion pulled left" ,
  ylim = c(0, 1) ,
  xaxt = "n" ,
  xlim = c(1, 4)
)
axis(1 ,
     at = 1:4 ,
     labels = c("R/N", "L/N", "R/P", "L/P"))
for (i in 1:100){
  lines(1:4 ,
        p_raw_asim[i, ] ,
        col = col.alpha("black", 0.25) ,
        lwd = 2)
}
```

