---
title: "08_03_2020_HW"
author: "John D."
date: "8/3/2020"
output:
  github_document:
    pandoc_args: --webtex
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rethinking)
library(tidyverse)
```

## 16M1.  Modify the cylinder height model, m16.1, so that the exponent 3 on height is instead a free parameter. Do you recover the value of 3 or not? Plot the posterior predictions for the new model. How do they differ from those of m16.1?

```{r}
## Load data
data(Howell1)
d <- Howell1
# scale observed variables
d$w <- d$weight / mean(d$weight)
d$h <- d$height / mean(d$height)
```

```{r}
## Original Model
m16.1a <- ulam(
  alist(
    w ~ dlnorm(mu , sigma),
    exp(mu) <- 3.141593 * k * p ^ 2 * h ^ 3,
    p ~ beta(2 , 18),
    k ~ exponential(0.5),
    sigma ~ exponential(1)
  ),
  data = d ,
  chains = 4 ,
  cores = 4,
  log_lik = T
)

precis(m16.1a)
```

```{r}
## New Model
m16.1b <- ulam(
  alist(
    w ~ dlnorm(mu , sigma),
    exp(mu) <- 3.141593 * k * p ^ 2 * h ^ Eh,
    p ~ beta(2 , 18),
    k ~ exponential(0.5),
    Eh ~ exponential(0.5),
    sigma ~ exponential(1)
  ),
  data = d ,
  chains = 4 ,
  cores = 4,
  log_lik = T
)

precis(m16.1b)
```

```{r}
## Another New Model
m16.1c <- ulam(
  alist(
    w ~ dlnorm(mu , sigma),
    exp(mu) <- 3.141593 * k * p ^ 2 * h ^ Eh,
    p ~ beta(2 , 18),
    k ~ exponential(0.5),
    Eh ~ exponential(1),
    sigma ~ exponential(1)
  ),
  data = d ,
  chains = 4 ,
  cores = 4,
  log_lik = T
)

precis(m16.1c)
```

```{r}
## Yet Another New Model
m16.1d <- ulam(
  alist(
    w ~ dlnorm(mu , sigma),
    exp(mu) <- 3.141593 * k * p ^ 2 * h ^ Eh,
    p ~ beta(2 , 18),
    k ~ exponential(0.5),
    Eh ~ normal(0, 2),
    sigma ~ exponential(1)
  ),
  data = d ,
  chains = 4 ,
  cores = 4,
  log_lik = T
)

precis(m16.1d)
```

```{r}
compare(m16.1a, m16.1b, m16.1c, m16.1d)
```

```{r}
pairs(m16.1a)
```
```{r}
pairs(m16.1b)
```

```{r}
par(mfrow = c(1,2))
h_seq <- seq( from=0 , to=max(d$h) , length.out=30 )
w_sim <- sim( m16.1a , data=list(h=h_seq) )
mu_mean <- apply( w_sim , 2 , mean )
w_CI <- apply( w_sim , 2 , PI )
plot( d$h , d$w , xlim=c(0,max(d$h)) , ylim=c(0,max(d$w)) , col=rangi2 ,
lwd=2 , xlab="height (scaled)" , ylab="weight (scaled)" )
lines( h_seq , mu_mean )
shade( w_CI , h_seq )

h_seq <- seq( from=0 , to=max(d$h) , length.out=30 )
w_sim <- sim( m16.1b , data=list(h=h_seq) )
mu_mean <- apply( w_sim , 2 , mean )
w_CI <- apply( w_sim , 2 , PI )
plot( d$h , d$w , xlim=c(0,max(d$h)) , ylim=c(0,max(d$w)) , col=rangi2 ,
lwd=2 , xlab="height (scaled)" , ylab="weight (scaled)" )
lines( h_seq , mu_mean )
shade( w_CI , h_seq )
```

Did not get 3 back but a value closer to 2.32. It also appears to handle the short values better and tall values worse. This model probably needs to factor in age or be splines for different height rangles.

## 16M2. Conduct a prior predictive simulation for the cylinder height model. Begin with the priors in the chapter. Do these produce reasonable prior height distributions? If not, which modifications do you suggest?

This part is coveredi n 16.3.2 of the book

```{r}
N <- 1e4
p <- rbeta(N, 2 , 18)
k <- rexp(N, 0.5)

# Prior relative height to weight curve
plot(
  d$h,
  d$w,
  xlim = c(0, max(d$h)) ,
  ylim = c(0, max(d$w)),
  xlab = "height (scaled)",
  ylab = "weight (scaled)"
)
for (i in 1:100)
  curve(
    exp(log(k[i]) + log(3.141593) + 2*log(p[i]) + 3*log(x)),
    add = TRUE ,
    col = grau() ,
    lwd = 1.5
    
  )
# Posterior relative height to weight curve
h_seq <- seq( from=0 , to=max(d$h) , length.out=30 )
w_sim <- sim( m16.1a , data=list(h=h_seq) )
mu_mean <- apply( w_sim , 2 , mean )
w_CI <- apply( w_sim , 2 , PI )
plot( d$h , d$w , xlim=c(0,max(d$h)) , ylim=c(0,max(d$w)) , col=rangi2 ,
lwd=2 , xlab="height (scaled)" , ylab="weight (scaled)" )
lines( h_seq , mu_mean )
shade( w_CI , h_seq )
```

Probably larger k

```{r}
N <- 1e4
p <- rbeta(N, 2 , 18)
k <- rexp(N, 0.1)

# Prior relative height to weight curve
plot(
  d$h,
  d$w,
  xlim = c(0, max(d$h)) ,
  ylim = c(0, max(d$w)),
  xlab = "height (scaled)",
  ylab = "weight (scaled)"
)
for (i in 1:100)
  curve(
    exp(log(k[i]) + log(3.141593) + 2*log(p[i]) + 3*log(x)),
    add = TRUE ,
    col = grau() ,
    lwd = 1.5
    
  )
# Posterior relative height to weight curve
h_seq <- seq( from=0 , to=max(d$h) , length.out=30 )
w_sim <- sim( m16.1a , data=list(h=h_seq) )
mu_mean <- apply( w_sim , 2 , mean )
w_CI <- apply( w_sim , 2 , PI )
plot( d$h , d$w , xlim=c(0,max(d$h)) , ylim=c(0,max(d$w)) , col=rangi2 ,
lwd=2 , xlab="height (scaled)" , ylab="weight (scaled)" )
lines( h_seq , mu_mean )
shade( w_CI , h_seq )
```