---
title: "08_03_2020_HW"
author: "John D."
date: "8/3/2020"
output:
  github_document:
    pandoc_args: --webtex
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rethinking)
library(tidyverse)
```

## 16M1.  Modify the cylinder height model, m16.1, so that the exponent 3 on height is instead a free parameter. Do you recover the value of 3 or not? Plot the posterior predictions for the new model. How do they differ from those of m16.1?

```{r}
## Load data
data(Howell1)
d <- Howell1
# scale observed variables
d$w <- d$weight / mean(d$weight)
d$h <- d$height / mean(d$height)
```

```{r}
## Original Model
m16.1 <- ulam(
  alist(
    w ~ dlnorm(mu , sigma),
    exp(mu) <- 3.141593 * k * p ^ 2 * h ^ 3,
    p ~ beta(2 , 18),
    k ~ exponential(0.5),
    sigma ~ exponential(1)
  ),
  data = d ,
  chains = 4 ,
  cores = 4
)

precis(m16.1)
```


## 16M2. Conduct a prior predictive simulation for the cylinder height model. Begin with the priors in the chapter. Do these produce reasonable prior height distributions? If not, which modifications do you suggest?
