---
title: "08_31_2020_HW"
author: "John D."
date: "08_31_2020"
output:
  github_document:
    pandoc_args: --webtex
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rethinking)
library(tidyverse)
library(mefa)
```

```{r}
dat <- read_csv("light_round1_tall.csv")
head(dat)
```

### 1. Subset the data to only include wps==0.  We may look at water potential in a future week.

```{r}
dat.small <- dat %>%
  filter(wps == 0)
head(dat.small)
```

### 2. Let's start with a single pop "STDI".  Evaluate models relate germination rate to temperature.  Explain your model choice and prior choice.  Try to do this both with ulam and brm.

```{r}
dat.smaller <- dat.small %>%
  filter(pops == "STDI")
head(dat.smaller)

ggplot(dat.smaller, aes(x = day, y = cumulative_germ, color = as.factor(temps))) +
  geom_line() +
  geom_point()

ggplot(dat.smaller, aes(x = day, y = cumulative_prop_germ, color = as.factor(temps))) +
  geom_line() +
  geom_point()
```

Following the cat model. Probability of no germination conditional on the observation time so far.

Need to make each seed a single event

```{r}
# Add germinated
germs.dat <- data.frame(matrix(ncol = 11, nrow = 0))
colnames(germs.dat) <- colnames(dat.smaller)
germs <- dat.smaller %>%
  filter(germ != 0)

for(i in 1:nrow(germs)){
  germs.dat <- rbind(germs.dat, rep(germs[i,], germs[[i,6]]))
}
germs.dat$germ <- 1

# Add ungerminated

no.germs.dat <- data.frame(matrix(ncol = 11, nrow = 0))
colnames(no.germs.dat) <- colnames(dat.smaller)

no.germs <- dat.smaller %>% group_by(temps) %>%
  summarize(germs = sum(germ),
            nogerm = max(total_seeds) - germs) %>%
  pull(nogerm)

nogerms <- dat.smaller %>% 
  group_by(temps) %>%
  filter(day == max(dat.smaller$day))

for(i in 1:nrow(nogerms)){
  no.germs.dat <- rbind(no.germs.dat, rep(nogerms[i,], no.germs[i]))
}
no.germs.dat$germ <- 0

new.dat <- rbind(germs.dat,no.germs.dat)
```

Group by temp, days will parameter

```{r}
dat_list <- list(
  Germ = new.dat$germ,
  Days = new.dat$day,
  Temp = as.numeric(as.factor(new.dat$temps))
)

model.1 <- ulam(
  alist(
    Days|Germ==1 ~ exponential(lambda),
    Days|Germ==0 ~ custom(exponential_lccdf( !Y | lambda)),
    lambda <- 1.0/mu,
    log(mu) <- a[Temp],
    a[Temp] ~ normal(0,1)
  ), data = dat_list, chains = 4, cores = 4
)
precis(model.1, depth = 2)

post <- extract.samples( model.1 )
post$D <- exp(post$a)
precis( post , 2 )

```

```{r}
days.seq <- rep(seq(from = 1, to = 28), 8)
temps.seq <- rep(seq(from = 1, to = 8), 28)
mu <- link(model.1, data = data.frame(Days = days.seq, Temp = temps.seq))
```

```{r}

```

