---
title: "Ch.7 Notes"
author: "John D."
date: "July 23, 2019"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(rethinking)
```

# 7 Ulysses’ Compass
## 7.1. The problem with parameters
### 7.1.1. More parameters always improve fit
```{r}
sppnames <- c( "afarensis","africanus","habilis","boisei",
"rudolfensis","ergaster","sapiens")
brainvolcc <- c( 438 , 452 , 612, 521, 752, 871, 1350 )
masskg <- c( 37.0 , 35.5 , 34.5 , 41.5 , 55.5 , 61.0 , 53.5 )
d <- data.frame( species=sppnames , brain=brainvolcc , mass=masskg )
d$mass_std <- (d$mass - mean(d$mass))/sd(d$mass)
d$brain_std <- d$brain / max(d$brain)
m7.1 <- quap(
  alist(
    brain_std ~ dnorm( mu , exp(log_sigma) ),
    mu <- a + b*mass_std,
    a ~ dnorm( 0.5 , 1 ),
    b ~ dnorm( 0 , 10 ),
    log_sigma ~ dnorm( 0 , 1 )
    ),
  data=d
  )
set.seed(12)
s <- sim( m7.1 )
r <- apply(s,2,mean) - d$brain_std
resid_var <- var2(r)
outcome_var <- var2( d$brain_std )
1 - resid_var/outcome_var
```
```{r}
R2_is_bad <- function( quap_fit ) {
  s <- sim( quap_fit , refresh=0 )
  r <- apply(s,2,mean) - d$brain_std
  1 - var2(r)/var2(d$brain_std)
}
```
```{r}
m7.2 <- quap(
  alist(
    brain_std ~ dnorm( mu , exp(log_sigma) ),
    mu <- a + b[1]*mass_std + b[2]*mass_std^2,
    a ~ dnorm( 0.5 , 1 ),
    b ~ dnorm( 0 , 10 ),
    log_sigma ~ dnorm( 0 , 1 )
    ),
  data=d,
  start=list(b=rep(0,2))
  )
```
```{r}
m7.3 <- quap(
  alist(
    brain_std ~ dnorm( mu , exp(log_sigma) ),
    mu <- a + b[1]*mass_std + b[2]*mass_std^2 +
      b[3]*mass_std^3,
    a ~ dnorm( 0.5 , 1 ),
    b ~ dnorm( 0 , 10 ),
    log_sigma ~ dnorm( 0 , 1 )
    ),
  data=d,
  start=list(b=rep(0,3))
  )
```
```{r}
m7.4 <- quap(
  alist(
    brain_std ~ dnorm( mu , exp(log_sigma) ),
    mu <- a + b[1]*mass_std + b[2]*mass_std^2 +
      b[3]*mass_std^3 + b[4]*mass_std^4,
    a ~ dnorm( 0.5 , 1 ),
    b ~ dnorm( 0 , 10 ),
    log_sigma ~ dnorm( 0 , 1 )
    ),
  data=d,
  start=list(b=rep(0,4))
  )
```
```{r}
m7.5 <- quap(
  alist(
    brain_std ~ dnorm( mu , exp(log_sigma) ),
    mu <- a + b[1]*mass_std + b[2]*mass_std^2 +
      b[3]*mass_std^3 + b[4]*mass_std^4 +
      b[5]*mass_std^5,
    a ~ dnorm( 0.5 , 1 ),
    b ~ dnorm( 0 , 10 ),
    log_sigma ~ dnorm( 0 , 1 )
    ),
  data=d,
  start=list(b=rep(0,5))
  )
```
```{r}
m7.6 <- quap(
  alist(
    brain_std ~ dnorm( mu , 0.001 ),
    mu <- a + b[1]*mass_std + b[2]*mass_std^2 +
      b[3]*mass_std^3 + b[4]*mass_std^4 +
      b[5]*mass_std^5 + b[6]*mass_std^6,
    a ~ dnorm( 0.5 , 1 ),
    b ~ dnorm( 0 , 10 )
    ),
  data=d,
  start=list(b=rep(0,6))
  )
```
```{r}
post <- extract.samples(m7.1)
mass_seq <- seq( from=min(d$mass_std) , to=max(d$mass_std) , length.out=100 )
l <- link( m7.1 , data=list( mass_std=mass_seq ) )
mu <- apply( l , 2 , mean )
ci <- apply( l , 2 , PI )
plot( brain_std ~ mass_std , data=d )
lines( mass_seq , mu )
shade( ci , mass_seq )
```
### 7.1.2. Too few parameters hurts, too
```{r}
m7.7 <- quap(
  alist(
    brain_std ~ dnorm( mu , exp(log_sigma) ),
    mu <- a,
    a ~ dnorm( 0.5 , 1 ),
    log_sigma ~ dnorm( 0 , 1 )
    ),
  data=d )
```
## 7.2. Entropy and accuracy
### 7.2.1. Firing the weatherperson.
#### 7.2.1.1 Costs and benefits
#### 7.2.1.2. Measuring accuracy
### 7.2.2. Information and uncertainty
```{r}
p <- c( 0.3 , 0.7 )
-sum( p*log(p) )
p <- c( 0.01 , 0.99 )
-sum( p*log(p) )
```
### 7.2.3. From entropy to accuracy
### 7.2.4. Estimating divergence
```{r}
set.seed(1)
lppd( m7.1 , n=1e4 )
```
### 7.2.5. Scoring the right data
```{r}
set.seed(1)
sapply( list(m7.1,m7.2,m7.3,m7.4,m7.5,m7.6) , function(m) sum(lppd(m)) )
```


```{r}
set.seed(1)
logprob <- sim( m7.1 , ll=TRUE , n=1e4 )
n <- ncol(logprob)
ns <- nrow(logprob)
f <- function( i ) log_sum_exp( logprob[,i] ) - log(ns)
( lppd <- sapply( 1:n , f ) )
set.seed(1)
sapply( list(m7.1,m7.2,m7.3,m7.4,m7.5,m7.6) , function(m) sum(lppd(m)) )
```


```{r}
# N <- 20
# kseq <- 1:5
# dev <- sapply( kseq , function(k) {
#   print(k);
#   r <- replicate( 1e4 , sim_train_test( N=N, k=k ) );
#   c( mean(r[1,]) , mean(r[2,]) , sd(r[1,]) , sd(r[2,]) )
# } )
# r <- mcreplicate( 1e4 , sim_train_test( N=N, k=k ) , mc.cores=1 )
# plot( 1:5 , dev[1,] , ylim=c( min(dev[1:2,])-5 , max(dev[1:2,])+10 ) ,
# xlim=c(1,5.1) , xlab="number of parameters" , ylab="deviance" ,
# pch=16 , col=rangi2 )
# mtext( concat( "N = ",N ) )
# points( (1:5)+0.1 , dev[2,] )
# for ( i in kseq ) {
# pts_in <- dev[1,i] + c(-1,+1)*dev[3,i]
# pts_out <- dev[2,i] + c(-1,+1)*dev[4,i]
# lines( c(i,i) , pts_in , col=rangi2 )
# lines( c(i,i)+0.1 , pts_out )
# }
```

## 7.3. Golem Taming: Regularization
## 7.4. Predicting predictive accuracy
### 7.4.1. Cross-validation
### 7.4.2. Information criteria
```{r}
data(cars)
m <- quap(
  alist(
    dist ~ dnorm(mu,sigma),
    mu <- a + b*speed,
    a ~ dnorm(0,100),
    b ~ dnorm(0,10),
    sigma ~ dexp(1)
    ),
  data=cars
  )
set.seed(94)
post <- extract.samples(m,n=1000)
```

```{r}
n_samples <- 1000
logprob <- sapply( 1:n_samples ,
function(s) {
mu <- post$a[s] + post$b[s]*cars$speed
dnorm( cars$dist , mu , post$sigma[s] , log=TRUE )
} )
```

```{r}
n_cases <- nrow(cars)
lppd <- sapply( 1:n_cases , function(i) log_sum_exp(logprob[i,]) - log(n_samples) )
```

```{r}
sum(lppd)
```

```{r}
pWAIC <- sapply( 1:n_cases , function(i) var(logprob[i,]) )
sum(pWAIC)
```

```{r}
-2*( sum(lppd) - sum(pWAIC) )
```

```{r}
WAIC(m)
```

```{r}
waic_vec <- -2*( lppd - pWAIC )
sqrt( n_cases*var(waic_vec) )
```

### 7.4.3. Comparing LOOCV, LOOIS, and WAIC

## 7.5. Using cross-validation and information criteria
### 7.5.1. Model mis-selection

```{r}
set.seed(71)
# number of plants
N <- 100
# simulate initial heights
h0 <- rnorm(N,10,2)
# assign treatments and simulate fungus and growth
treatment <- rep( 0:1 , each=N/2 )
fungus <- rbinom( N , size=1 , prob=0.5 - treatment*0.4 )
h1 <- h0 + rnorm(N, 5 - 3*fungus)
# compose a clean data frame
d <- data.frame( h0=h0 , h1=h1 , treatment=treatment , fungus=fungus )
```

```{r}
m6.6 <- quap(
  alist(
    h1 ~ dnorm( mu , sigma ),
    mu <- h0*p,
    p ~ dlnorm( 0 , 0.25 ),
    sigma ~ dexp( 1 )
    ),
  data=d )
```
```{r}
m6.7 <- quap(
  alist(
    h1 ~ dnorm( mu , sigma ),
    mu <- h0 * p,
    p <- a + bt*treatment + bf*fungus,
    a ~ dlnorm( 0 , 0.2 ) ,
    bt ~ dnorm( 0 , 0.5 ),
    bf ~ dnorm( 0 , 0.5 ),
    sigma ~ dexp( 1 )
    ),
  data=d )
```

```{r}
m6.8 <- quap(
  alist(
    h1 ~ dnorm( mu , sigma ),
    mu <- h0 * p,
    p <- a + bt*treatment,
    a ~ dlnorm( 0 , 0.2 ),
    bt ~ dnorm( 0 , 0.5 ),
    sigma ~ dexp( 1 )
    ),
  data=d )
```

```{r}
set.seed(11)
WAIC( m6.7 )
```

```{r}
set.seed(77)
compare( m6.6 , m6.7 , m6.8 )
```

```{r}
set.seed(91)
waic_m6.6 <- WAIC( m6.6 , pointwise=TRUE )
waic_m6.7 <- WAIC( m6.7 , pointwise=TRUE )
waic_m6.8 <- WAIC( m6.8 , pointwise=TRUE )
n <- length(waic_m6.6)
diff_m6.7_m6.8 <- waic_m6.7 - waic_m6.8
diff_m6.7_m6.8
```

```{r}
40.0 + c(-1,1)*10.4*2.6
```

```{r}
plot( compare( m6.6 , m6.7 , m6.8 ) )
```

```{r}
set.seed(92)
waic_m6.6 <- WAIC( m6.6 , pointwise=TRUE )
diff_m6.6_m6.8 <- waic_m6.6 - waic_m6.8
sqrt( n*var( diff_m6.6_m6.8 ) )
```

```{r}
set.seed(93)
compare( m6.6 , m6.7 , m6.8 )@dSE
```

### 7.5.2. Something about Cebus

```{r}
data(Primates301)
d <- Primates301
d$log_L <- scale( log(d$longevity) )
d$log_B <- scale( log(d$brain) )
d$log_M <- scale( log(d$body) )
sapply( d[,c("log_L","log_B","log_M")] , function(x) sum(is.na(x)) )
d2 <- d[ complete.cases( d$log_L , d$log_M , d$log_B ) , ]
nrow(d2)
```

```{r}
m7.8 <- quap(
  alist(
    log_L ~ dnorm( mu , sigma ),
    mu <- a + bM*log_M + bB*log_B,
    a ~ dnorm(0,0.1),
    bM ~ dnorm(0,0.5),
    bB ~ dnorm(0,0.5),
    sigma ~ dexp(1)),
  data=d2 )
```

```{r}
m7.9 <- quap(
  alist(
    log_L ~ dnorm( mu , sigma ),
    mu <- a + bB*log_B,
    a ~ dnorm(0,0.1),
    bB ~ dnorm(0,0.5),
    sigma ~ dexp(1)),
  data=d2 )
```

```{r}
m7.10 <- quap(
  alist(
    log_L ~ dnorm( mu , sigma ),
    mu <- a + bM*log_M,
    a ~ dnorm(0,0.1),
    bM ~ dnorm(0,0.5),
    sigma ~ dexp(1)),
  data=d2 )
```

```{r}
set.seed(301)
compare( m7.8 , m7.9 , m7.10 )
```

```{r}
plot( compare( m7.8 , m7.9 , m7.10 ) )
```

```{r}
plot( coeftab( m7.8 , m7.9 , m7.10 ) , pars=c("bM","bB") )
```

```{r}
cor( d2$log_B , d2$log_M )
```

```{r}
waic_m7.8 <- WAIC( m7.8 , pointwise=TRUE )
waic_m7.9 <- WAIC( m7.9 , pointwise=TRUE )
```

```{r}
# compute point scaling
x <- d2$log_B - d2$log_M
x <- x - min(x)
x <- x / max(x)
# draw the plot
plot( waic_m7.8 - waic_m7.9 , d2$log_L ,
xlab="pointwise difference in WAIC" , ylab="log longevity (std)" , pch=21 ,
col=col.alpha("black",0.8) , cex=1+x , lwd=2 , bg=col.alpha(rangi2,0.4) )
abline( v=0 , lty=2 )
abline( h=0 , lty=2 )
```

```{r}
m7.11 <- quap(
  alist(
    log_B ~ dnorm( mu , sigma ),
    mu <- a + bM*log_M + bL*log_L,
    a ~ dnorm(0,0.1),
    bM ~ dnorm(0,0.5),
    bL ~ dnorm(0,0.5),
    sigma ~ dexp(1)),
  data=d2 )
precis( m7.11 )
```

```{r}
m7.12 <- quap(
  alist(
    log_B ~ dnorm( mu , sigma ),
    mu <- a + bM*log_M,
    a ~ dnorm(0,0.1),
    bM ~ dnorm(0,0.5),
    sigma ~ dexp(1)),
  data=d2 )
precis( m7.12 )
```

```{r}
m7.13 <- quap(
  alist(
    log_B ~ dnorm( mu , sigma ),
    mu <- a + bL*log_L,
    a ~ dnorm(0,0.1),
    bL ~ dnorm(0,0.5),
    sigma ~ dexp(1)),
  data=d2 )
precis( m7.13 )
```