---
title: "12_6_2019_HW"
author: "John D."
date: "12/6/2019"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(rethinking)
options(mc.cores = parallel::detectCores())
rstan_options(auto_write = TRUE)
Sys.setenv(LOCAL_CPPFLAGS = '-march=native')
```

# Book

## 12E1. What is the difference between an ordered categorical variable and an unordered one? Define and then give an example of each.

An ordered categorical variable contains structure where one value is greater than the other. An example might be gauging the temperature of an object using (cold, cold, warm, hot) instead of the actual temperature. An unordered catergorical variable is one where there is no structure. An example would be a variable of favorite color. No color is ordered higher than the other.

## 12E2. What kind of link function does an ordered logistic regression employ? How does it differ from an ordinary logit link?

An ordered logistic regression employs the cumulative logit where it is the odds of witnessing the event and any event smaller than it in the ordering.

## 12M1. At a certain university, employees are annually rated from 1 to 4 on their productivity, with 1 being least productive and 4 most productive. In a certain department at this certain university in a certain year, the numbers of employees receiving each rating were (from 1 to 4): 12, 36, 7, 41. Compute the log cumulative odds of each rating

```{r}
ratings <- c(12,36,7,41)

ratings_por <- ratings  / sum(ratings)
ratings_por

cum_ratings <- cumsum(ratings_por)
cum_ratings

round( lco <- logit( cum_ratings ) , 2 )
names(lco) <- c("1","2","3","4")
lco
```

## 12M2. Make a version of Figure 12.5 for the employee ratings data given just above.

```{r}
cum_ratings
dat <- data.frame(rating = 1:4, cum_prob = cum_ratings, probs = ratings_por)
ggplot(dat, aes(x = rating, y = cum_ratings)) +
  geom_linerange(aes(ymin = 0, ymax = cum_prob)) +
  geom_linerange(aes(x = rating + 0.025,
                     ymin = cum_prob - probs,
                     ymax = cum_prob),
                 color = "blue") +
  geom_line() +
  geom_point(shape = 1) +
  geom_text(aes(y = (2*cum_prob - probs)/2,
                label = 1:4),
            color="blue",
            position = position_nudge(x=.15)) +
  ylab("cumulative proportion")
```

# PDF

## 1. In the Trolley data--`data(Trolley)`-- we saw how education level (modeled as an ordered category) is associate with responses. Is this association causal? One plausible confound is that education is also associated with age, through a causal process: People are older when they finish school than when they begin it.
## Reconsider the Trolley data in this light. Draw a DAG that represents hypothetical causal relationships among response, education, and age. Which statistical model or models do you need to evauluate the causal influence of education on responses? Fit these models to the trolley data. What do you conclude about the causal relationships among these three variables?

```{r}
library(dagitty)
# Build DAG
dag <- dagitty( "dag {
                A -> R
                E -> R
                A -> E
                }")
coordinates(dag) <- list(x = c(A=0,R=0.5,E=1), y = c(A=0,R=-0.5,E=0))
plot(dag)

data("Trolley")
d <- Trolley

edu_levels <- c( 6 , 1 , 8 , 4 , 7 , 2 , 5 , 3 )
d$edu_new <- as.integer(edu_levels[ d$edu ])

dat <- list(
  R = d$response ,
  action = d$action,
  intention = d$intention,
  contact = d$contact,
  E = d$edu_new,
  alpha = rep(2 , 7),
  age = standardize(d$age)
)

mdl_1 <- ulam(
  alist(
    R ~ ordered_logistic(phi , kappa),
    phi <- bE * sum(delta_j[1:E]) + bA * action + bI * intention + bC * contact + bAge * age,
    kappa ~ normal(0 , 1.5),
    c(bA, bI, bC, bE, bAge) ~ normal(0 , 1),
    vector[8]:delta_j <<- append_row(0 , delta),
    simplex[7]:delta ~ dirichlet(alpha)
  ),
  data = dat ,
  chains = 4 ,
  cores = 4 ,
  coerce_int = FALSE
)

precis(mdl_1, depth = 2)
```

Education now leads to a more permissable result while age leads to less permissable. Back door may be real

## 2. Consider one more variable in the Trolley data: Gender. Suppose that gender might influence education as well as response directly. Draw a DAG now that includes response, education, age, and gender.
## Using only the DAG, is it possible that the inferences from Problem 1 are confounded by gender? If so, define any additional models you need to infer the causal influence of education on response. What do you conclude?

```{r}
# Build DAG
dag <- dagitty( "dag {
                A -> R
                E -> R
                A -> E
                G -> R
                G -> E
                }")
coordinates(dag) <- list(x = c(A=0,R=0.5,E=0.5,G=1), y = c(A=0,R=-0.5,E=0,G=0))
plot(dag)

dat <- list(
  R = d$response ,
  action = d$action,
  intention = d$intention,
  contact = d$contact,
  E = d$edu_new,
  alpha = rep(2 , 7),
  age = standardize(d$age),
  male = d$male
)

mdl_2 <- ulam(
  alist(
    R ~ ordered_logistic(phi , kappa),
    phi <- bE * sum(delta_j[1:E]) + bA * action + bI * intention + bC * contact + bAge * age + bM * male,
    kappa ~ normal(0 , 1.5),
    c(bA, bI, bC, bE, bAge, bM) ~ normal(0 , 1),
    vector[8]:delta_j <<- append_row(0 , delta),
    simplex[7]:delta ~ dirichlet(alpha)
  ),
  data = dat ,
  chains = 4 ,
  cores = 4 ,
  coerce_int = FALSE
)

precis(mdl_2, depth = 2)
```

Education does not play a role anymore. Now both gender and age play the major role with the elderly and females providing less permissable answers.