---
title: "Notes"
author: "John D."
date: "08_31_2020"
output:
  github_document:
    pandoc_args: --webtex
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rethinking)
library(tidyverse)
```

# 12 Monsters and Mixtures
## 12.1. Over-dispersed counts
### 12.1.1. Beta-binomial
#### R code 12.1
```{r}
pbar <- 0.5
theta <- 5
curve( dbeta2(x,pbar,theta) , from=0 , to=1 ,
xlab="probability" , ylab="Density" )
pbar <- 0.5
theta <- 2
curve( dbeta2(x,pbar,theta) , from=0 , to=1 ,
xlab="probability" , ylab="Density", add = T, col = "blue" )
pbar <- 0.5
theta <- 1
curve( dbeta2(x,pbar,theta) , from=0 , to=1 ,
xlab="probability" , ylab="Density", add = T, col = "red" )
```

```{r}
pbar <- 0.2
theta <- 5
curve( dbeta2(x,pbar,theta) , from=0 , to=1 ,
xlab="probability" , ylab="Density" )
pbar <- 0.2
theta <- 2
curve( dbeta2(x,pbar,theta) , from=0 , to=1 ,
xlab="probability" , ylab="Density", add = T, col = "blue" )
pbar <- 0.2
theta <- 1
curve( dbeta2(x,pbar,theta) , from=0 , to=1 ,
xlab="probability" , ylab="Density", add = T, col = "red" )
```

#### R code 12.2
```{r}
library(rethinking)
data(UCBadmit)
d <- UCBadmit
d$gid <- ifelse(d$applicant.gender == "male" , 1L , 2L)
dat <- list(A = d$admit ,
            N = d$applications ,
            gid = d$gid)
m12.1 <- ulam(
  alist(
    A ~ dbetabinom(N , pbar , theta),
    logit(pbar) <- a[gid],
    a[gid] ~ dnorm(0 , 1.5),
    transpars > theta <<- phi + 2.0,
    phi ~ dexp(1
    )
  ), data=dat , chains=4 )
```

#### R code 12.3
```{r}
post <- extract.samples( m12.1 )
post$da <- post$a[,1] - post$a[,2]
precis( post , depth=2 )
```

```{r}
post$dp <- inv_logit(post$a[,1]) - inv_logit(post$a[,2])
precis( post , depth=2 )
```

```{r}
stancode(m12.1)
```

#### R code 12.4
```{r}
gid <- 2
# draw posterior mean beta distribution
curve(
  dbeta2(x, mean(logistic(post$a[, gid])), mean(post$theta)) ,
  from = 0 ,
  to = 1 ,
  ylab = "Density" ,
  xlab = "probability admit",
  ylim = c(0, 3) ,
  lwd = 2
)
# draw 50 beta distributions sampled from posterior
for (i in 1:50) {
  p <- logistic(post$a[i, gid])
  theta <- post$theta[i]
  curve(dbeta2(x, p, theta) ,
        add = TRUE ,
        col = col.alpha("black", 0.2))
}
mtext("distribution of female admission rates")
```

#### R code 12.5
```{r}
postcheck( m12.1 )
```

```{r}
dat_list$dept_id <- rep(1:6, each = 2)
m12.1a <- ulam(
  alist(
    A ~ dbetabinom(N , pbar , theta),
    logit(pbar) <- a[gid] + + delta[dept_id],
    a[gid] ~ dnorm(0 , 1.5),
    delta[dept_id] ~ dnorm(0 , 1.5),
    transpars > theta <<- phi + 2.0,
    phi ~ dexp(1
    )
  ), data=dat , chains=4 )

post <- extract.samples( m12.1a )
post$da <- post$a[,1] - post$a[,2]
precis( post , depth=2 )
```

```{r}
gid <- 2
# draw posterior mean beta distribution
curve(
  dbeta2(x, mean(logistic(post$a[, gid])), mean(post$theta)) ,
  from = 0 ,
  to = 1 ,
  ylab = "Density" ,
  xlab = "probability admit",
  ylim = c(0, 3) ,
  lwd = 2
)
# draw 50 beta distributions sampled from posterior
for (i in 1:50) {
  p <- logistic(post$a[i, gid])
  theta <- post$theta[i]
  curve(dbeta2(x, p, theta) ,
        add = TRUE ,
        col = col.alpha("black", 0.2))
}
mtext("distribution of female admission rates")
```

```{r}
postcheck(m12.1a)
```

### 12.1.2. Negative-binomial or gamma-Poisson
#### R code 12.6
```{r}
library(rethinking)
data(Kline)
d <- Kline
d$P <- standardize(log(d$population))
d$contact_id <- ifelse(d$contact == "high" , 2L , 1L)
dat2 <- list(
  T = d$total_tools,
  P = d$population,
  cid = d$contact_id
)
m12.2 <- ulam(
  alist(
    T ~ dgampois(lambda , phi),
    lambda <- exp(a[cid]) * P ^ b[cid] / g,
    a[cid] ~ dnorm(1, 1),
    b[cid] ~ dexp(1),
    g ~ dexp(1),
    phi ~ dexp(1)
  ),
  data = dat2 ,
  chains = 4 ,
  log_lik = TRUE
)
```

### 12.1.3. Over-dispersion, entropy, and information criteria
## 12.2. Zero-inflated outcomes
### 12.2.1. Example: Zero-inflated Poisson
#### R code 12.7
```{r}
# define parameters
prob_drink <- 0.2 # 20% of days
rate_work <- 1 # average 1 manuscript per day
# sample one year of production
N <- 365
# simulate days monks drink
set.seed(365)
drink <- rbinom(N , 1 , prob_drink)
# simulate manuscripts completed
y <- (1 - drink) * rpois(N , rate_work)
```

#### R code 12.8
```{r}
simplehist( y , xlab="manuscripts completed" , lwd=4 )
zeros_drink <- sum(drink)
zeros_work <- sum(y==0 & drink==0)
zeros_total <- sum(y==0)
lines( c(0,0) , c(zeros_work,zeros_total) , lwd=4 , col=rangi2 )
```

#### R code 12.9
```{r}
m12.3 <- ulam(
  alist(
    y ~ dzipois(p , lambda),
    logit(p) <- ap,
    log(lambda) <- al,
    ap ~ dnorm(-1.5 , 1),
    al ~ dnorm(1 , 0.5)
  ) ,
  data = list(y = y) ,
  chains = 4
)
precis(m12.3)
```

```{r}
stancode(m12.3)
```

#### R code 12.10
```{r}
post <- extract.samples( m12.3 )
mean( inv_logit( post$ap ) ) # probability drink
mean( exp( post$al ) ) # rate finish manuscripts, when not drinking
```

#### R code 12.11
```{r}
m12.3_alt <- ulam(
  alist(
    y | y > 0 ~ custom(log1m(p) + poisson_lpmf(y | lambda)),
    y | y == 0 ~ custom(log_mix(p , 0 , poisson_lpmf(0 | lambda))),
    logit(p) <- ap,
    log(lambda) <- al,
    ap ~ dnorm(-1.5, 1),
    al ~ dnorm(1, 0.5)
  ) ,
  data = list(y = as.integer(y)) ,
  chains = 4
)
precis(m12.3_alt)
post <- extract.samples( m12.3_alt )
mean( inv_logit( post$ap ) ) # probability drink
mean( exp( post$al ) ) # rate finish manuscripts, when not drinking
```

## 12.3. Ordered categorical outcomes
### 12.3.1. Example: Moral intuition
#### R code 12.12
```{r}
library(rethinking)
data(Trolley)
d <- Trolley
head(d)
```

### 12.3.2. Describing an ordered distribution with intercepts
#### R code 12.13
```{r}
simplehist( d$response , xlim=c(1,7) , xlab="response" )
```

#### R code 12.14
```{r}
# discrete proportion of each response value
pr_k <- table(d$response) / nrow(d)
# cumsum converts to cumulative proportions
cum_pr_k <- cumsum(pr_k)
# plot
plot(
  1:7 ,
  cum_pr_k ,
  type = "b" ,
  xlab = "response" ,
  ylab = "cumulative proportion" ,
  ylim = c(0, 1)
)
```

#### R code 12.15
```{r}
logit <- function(x) log(x/(1-x)) # convenience function
round( lco <- logit( cum_pr_k ) , 2 )
```

```{r}
plot(
  1:7 ,
  lco ,
  type = "b" ,
  xlab = "response" ,
  ylab = "log-cumulative-odds" ,
  ylim = c(-2, 2)
)
```

#### R code 12.16
```{r}
m12.4 <- ulam(
  alist(R ~ dordlogit(0 , cutpoints),
        cutpoints ~ dnorm(0 , 1.5)) ,
  data = list(R = d$response),
  chains = 4 ,
  cores = 4
)
```

#### R code 12.17
```{r}
m12.4q <- quap(
  alist(response ~ dordlogit(0 , c(a1, a2, a3, a4, a5, a6)),
        c(a1, a2, a3, a4, a5, a6) ~ dnorm(0 , 1.5)) ,
  data = d ,
  start = list(
    a1 = -2,
    a2 = -1,
    a3 = 0,
    a4 = 1,
    a5 = 2,
    a6 = 2.5
  )
)
```

#### R code 12.18
```{r}
precis( m12.4 , depth=2 )
```


#### R code 12.19
```{r}
round( inv_logit(coef(m12.4)) , 3 )
```

### 12.3.3. Adding predictor variables
#### R code 12.20
```{r}
round( pk <- dordlogit( 1:7 , 0 , coef(m12.4) ) , 2 )
```

#### R code 12.21
```{r}
sum( pk*(1:7) )
```

#### R code 12.22
```{r}
round( pk <- dordlogit( 1:7 , 0 , coef(m12.4)-0.5 ) , 2 )
```

#### R code 12.23
```{r}
sum( pk*(1:7) )
```

#### R code 12.24
```{r}
dat <- list(
  R = d$response,
  A = d$action,
  I = d$intention,
  C = d$contact
)
m12.5 <- ulam(
  alist(
    R ~ dordlogit(phi , cutpoints),
    phi <- bA * A + bC * C + BI * I ,
    BI <- bI + bIA * A + bIC * C ,
    c(bA, bI, bC, bIA, bIC) ~ dnorm(0 , 0.5),
    cutpoints ~ dnorm(0 , 1.5)
  ) ,
  data = dat ,
  chains = 4 ,
  cores = 4
)
precis(m12.5)
```

```{r}
stancode(m12.5)
```

#### R code 12.25
```{r}
plot( precis(m12.5) , xlim=c(-1.4,0) )
```

#### R code 12.26, 12.27, and 12.28
```{r}
plot(
  NULL ,
  type = "n" ,
  xlab = "intention" ,
  ylab = "probability" ,
  xlim = c(0, 1) ,
  ylim = c(0, 1) ,
  xaxp = c(0, 1, 1) ,
  yaxp = c(0, 1, 2)
)

kA <- 0 # value for action
kC <- 0 # value for contact
kI <- 0:1 # values of intention to calculate over
pdat <- data.frame(A=kA,C=kC,I=kI)
phi <- link( m12.5 , data=pdat )$phi

post <- extract.samples( m12.5 )
for ( s in 1:50 ) {
pk <- pordlogit( 1:6 , phi[s,] , post$cutpoints[s,] )
for ( i in 1:6 ) lines( kI , pk[,i] , col=grau(0.1) )
}
```

#### R code 12.29
```{r}
kA <- 0 # value for action
kC <- 1 # value for contact
kI <- 0:1 # values of intention to calculate over
pdat <- data.frame(A = kA, C = kC, I = kI)
s <- sim(m12.5 , data = pdat)
simplehist(s , xlab = "response")
```

## 12.4. Ordered categorical predictors
#### R code 12.30
```{r}
library(rethinking)
data(Trolley)
d <- Trolley
levels(d$edu)
```

#### R code 12.31
```{r}
edu_levels <- c( 6 , 1 , 8 , 4 , 7 , 2 , 5 , 3 )
d$edu_new <- edu_levels[ d$edu ]
```

#### R code 12.32
```{r}
library(gtools)
set.seed(1805)
delta <- rdirichlet( 10 , alpha=rep(2,7) )
str(delta)
```

#### R code 12.33
```{r}
h <- 3
plot(
  NULL ,
  xlim = c(1, 7) ,
  ylim = c(0, 0.4) ,
  xlab = "index" ,
  ylab = "probability"
)
for (i in 1:nrow(delta))
  lines(
    1:7 ,
    delta[i, ] ,
    type = "b" ,
    pch = ifelse(i == h, 16, 1) ,
    lwd = ifelse(i == h, 4, 1.5) ,
    col = ifelse(i == h, "black", col.alpha("black", 0.7))
  )
```

#### R code 12.34
```{r}
dat <- list(
  R = d$response ,
  action = d$action,
  intention = d$intention,
  contact = d$contact,
  E = as.integer(d$edu_new),
  # edu_new as an index
  alpha = rep(2 , 7)
) # delta prior

m12.6 <- ulam(
  alist(
    R ~ ordered_logistic(phi , kappa),
    phi <-
      bE * sum(delta_j[1:E]) + bA * action + bI * intention + bC * contact,
    kappa ~ normal(0 , 1.5),
    c(bA, bI, bC, bE) ~ normal(0 , 1),
    vector[8]:delta_j <<- append_row(0 , delta),
    simplex[7]:delta ~ dirichlet(alpha)
  ),
  data = dat ,
  chains = 4 ,
  cores = 4 ,
  coerce_int = FALSE
)
```

```{r}
stancode(m12.6)
```

#### R code 12.35
```{r}
precis( m12.6 , depth=2 , omit="kappa" )
```

#### R code 12.36
```{r}
delta_labels <- c("Elem","MidSch","SHS","HSG","SCol","Bach","Mast","Grad")
pairs( m12.6 , pars="delta" , labels=delta_labels )
```

#### R code 12.37
```{r}
dat$edu_norm <- normalize(d$edu_new)
m12.7 <- ulam(
  alist(
    R ~ ordered_logistic(mu , cutpoints),
    mu <- bE * edu_norm + bA * action + bI * intention + bC * contact,
    c(bA, bI, bC, bE) ~ normal(0 , 1),
    cutpoints ~ normal(0 , 1.5)
  ),
  data = dat ,
  chains = 4 ,
  cores = 4
)
precis(m12.7)
```

