---
title: "Growth Curve Models"
author: "John D."
date: "10_28_2020"
output:
  github_document:
    pandoc_args: --webtex
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
     

```{r}
library(rethinking)
library(brms)
library(tidyverse)
```

Work on the attached data set from Maloof (2001).  It is a series of growth response curves, either in response to different light or hormone treatments.  The question you want to answer is whether there are differences in response slopes between the genotypes.  Note that the responses tend to be ~ log-linear with respect to the treatment.  I would recommend picking one of the treatments that approximates (log) linear such as blue, red, or GA.  You may want to truncate the data to only retain the linear part. But if you want to get fancy don't truncate and/or try some of the more challenging treatments (e.g. white).

## Starting with red light
```{r}
red <- read_csv("Red.csv")
head(red)
```
```{r}
ggplot(data = red, aes(x = length, color = as.factor(concentration))) +
  geom_density() +
  facet_wrap(~genotype)
```

```{r}
summary(standardize(log(red$length)))
```


```{r}
red.small <-  red[complete.cases(red$length), ]
dat_list <- list(
  L = standardize(log(red.small$length)),
  G = as.factor(red.small$genotype),
  C = red.small$concentration
)

model_1 <- ulam(
  alist(
    L ~ dnorm(mu, sigma),
    mu <- a[G] + bC[G]*C,
    a[G] ~ dnorm(0, 1),
    bC[G] ~ dnorm(0, 1),
    sigma ~ dexp(1)
  ) ,
  data = dat_list,
  chains = 4 ,
  cores = 4
)
precis(model_1, depth = 2)
```
```{r}
traceplot(model_1)
```

```{r}
trankplot(model_1)
```

```{r}
pairs(model_1)
```
# Plotting
```{r}
red.aves <- red.small %>%
  mutate(length = standardize(log(length))) %>%
  group_by(genotype, concentration) %>%
  mutate(ave_length = mean(length)) %>%
  ungroup() %>%
  select(-length) %>%
  unique() %>%
  arrange(genotype)

pred_data <- data.frame(G = rep(as.factor(unique(red$genotype)), each = 8),
                        C = rep(unique(red$concentration), 6))

mu <- link(model_1, data=pred_data,)
mu.mean <- apply(mu , 2, mean)
pred_data$mu <- mu.mean
mu.PI <- apply(mu, 2, PI, prob=0.89)
pred_data$Low <- mu.PI[1,]
pred_data$High <- mu.PI[2,]
pred_data$Original <- red.aves$ave_length


ggplot(data = pred_data,
       aes(x = C, y = mu, color = as.factor(G))) +
  geom_line(lty = 1) +
  geom_line(aes(x = C, y = Original), lty = 2)
```

## No standardization
```{r}
dat_list <- list(
  L = red.small$length,
  G = as.factor(red.small$genotype),
  C_1 = red.small$concentration,
  C_2 = red.small$concentration^2
)

model_2 <- ulam(
  alist(
    L ~ dnorm(mu, sigma),
    mu <- a[G] + bC1[G]*C_1 + bC2[G]*C_2,
    a[G] ~ dnorm(8, 3),
    bC1[G] ~ dnorm(0, 4),
    bC2[G] ~ dnorm(0, 4),
    sigma ~ dexp(1)
  ) ,
  data = dat_list,
  chains = 4 ,
  cores = 4
)
precis(model_2, depth = 2)
```

```{r}
red.aves <- red.small %>%
  group_by(genotype, concentration) %>%
  mutate(ave_length = mean(length)) %>%
  ungroup() %>%
  select(-length) %>%
  unique() %>%
  arrange(genotype)

pred_data <- data.frame(G = rep(as.factor(unique(red$genotype)), each = 8),
                        C_1 = rep(unique(red$concentration), 6),
                        C_2 = rep(unique(red$concentration^2), 6))

mu <- link(model_2, data=pred_data,)
mu.mean <- apply(mu , 2, mean)
pred_data$mu <- mu.mean
mu.PI <- apply(mu, 2, PI, prob=0.89)
pred_data$Low <- mu.PI[1,]
pred_data$High <- mu.PI[2,]
pred_data$Original <- red.aves$ave_length


ggplot(data = pred_data,
       aes(x = C_1, y = mu, color = as.factor(G))) +
  geom_line(lty = 1) +
  geom_line(aes(x = C_1, y = Original), lty = 2)
```

## Back to basics, focus on only one genotype
```{r}
red.Ber <- red.small %>%
  filter(genotype == "Ber-0")

dat_list <- list(
  C = red.Ber$concentration,
  L = red.Ber$length
)
summary(red.Ber)
```

```{r}
model_3 <- ulam(
  alist(
    L ~ dnorm(mu, sigma),
    mu <- a + bC*C,
    a ~ dnorm(8,4),
    bC ~ dlnorm(0,1),
    sigma ~ dexp(1)
  ), data = dat_list, cores = 4, chains = 4
)

precis(model_3)
```

```{r}
red.aves <- red.Ber %>%
  group_by(genotype, concentration) %>%
  mutate(length = mean(length)) %>%
  ungroup() %>%
  unique() %>%
  arrange(genotype)


plt <- ggplot(data = red.aves, aes(x = concentration, y = length)) +
  geom_line(lty = 1)
plt




pred_data <- data.frame(C = red.aves$concentration)

mu <- link(model_3, data=pred_data,)
mu.mean <- apply(mu , 2, mean)
red.aves$predicted <- mu.mean

plt <- plt + geom_line(data = red.aves, aes(x = concentration, y = predicted), lty = 2)
plt
```

```{r}
model_4 <- ulam(
  alist(
    L ~ dnorm(mu, sigma),
    mu <- a - bC*C,
    a ~ dnorm(8,4),
    bC ~ dlnorm(0,1),
    sigma ~ dexp(1)
  ), data = dat_list, cores = 4, chains = 4
)

precis(model_4)
```

```{r}
red.aves <- red.Ber %>%
  group_by(genotype, concentration) %>%
  mutate(length = mean(length)) %>%
  ungroup() %>%
  unique() %>%
  arrange(genotype)


plt <- ggplot(data = red.aves, aes(x = concentration, y = length)) +
  geom_line(lty = 1)
plt




pred_data <- data.frame(C = red.aves$concentration)

mu <- link(model_4, data=pred_data,)
mu.mean <- apply(mu , 2, mean)
red.aves$predicted <- mu.mean

plt <- plt + geom_line(data = red.aves, aes(x = concentration, y = predicted), lty = 2)
plt
```

```{r}
dat_list$C_2 <- red.Ber$concentration^2
model_5 <- ulam(
  alist(
    L ~ dnorm(mu, sigma),
    mu <- a - bC*C - bC_2*C_2,
    a ~ dnorm(8,4),
    bC ~ dnorm(0,5),
    bC_2 ~ dnorm(0,5),
    sigma ~ dexp(1)
  ), data = dat_list, cores = 4, chains = 4
)

precis(model_5)
```

```{r}
red.aves <- red.Ber %>%
  group_by(genotype, concentration) %>%
  mutate(length = mean(length)) %>%
  ungroup() %>%
  unique() %>%
  arrange(genotype)


plt <- ggplot(data = red.aves, aes(x = concentration, y = length)) +
  geom_line(lty = 1)
plt

pred_data <- data.frame(C = red.aves$concentration,
                        C_2 = red.aves$concentration^2)

mu <- link(model_5, data=pred_data,)
mu.mean <- apply(mu , 2, mean)
red.aves$predicted <- mu.mean

plt <- plt + geom_line(data = red.aves, aes(x = concentration, y = predicted), lty = 2)
plt
```

```{r}
dat_list$C_2 <- red.Ber$concentration^2
model_6 <- ulam(
  alist(
    L ~ dnorm(mu, sigma),
    mu <- a - bC*C - bC_2*C_2,
    a ~ dnorm(8,4),
    c(bC, bC_2) ~ dexp(2),
    sigma ~ dexp(1)
  ), data = dat_list, cores = 4, chains = 4
)

precis(model_6)
```

```{r}
red.aves <- red.Ber %>%
  group_by(genotype, concentration) %>%
  mutate(length = mean(length)) %>%
  ungroup() %>%
  unique() %>%
  arrange(genotype)


plt <- ggplot(data = red.aves, aes(x = concentration, y = length)) +
  geom_line(lty = 1)
plt

pred_data <- data.frame(C = red.aves$concentration,
                        C_2 = red.aves$concentration^2)

mu <- link(model_6, data=pred_data,)
mu.mean <- apply(mu , 2, mean)
red.aves$predicted <- mu.mean

plt <- plt + geom_line(data = red.aves, aes(x = concentration, y = predicted), lty = 2)
plt
```

# Try brms

```{r}
model_7 <- brm(length ~ 0 + concentration,
               data = red.Ber,
               cores = 4)
summary(model_7)
```

```{r}
preds <- predict(model_7, newdata = data.frame(concentration = red.aves$concentration))

red.aves$predicted <- preds[,1]
plt <- ggplot(data = red.aves, aes(x = concentration, y = length)) +
  geom_line(lty = 1)
plt
plt <- plt + geom_line(aes(x = concentration, y = predicted), lty = 2)
plt
```

