---
title: "07_06_2020_HW"
author: "John D."
date: "7/06/2020"
output:
  github_document:
    pandoc_args: --webtex
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(rethinking)
library(tidyverse)
library(ape)
library(ade4)
```

## 1. Load the data as follows:
```{r}
data("carni70")
tre <- read.tree(text=carni70$tre)
d <- carni70$tab
```

```{r}
head(d)
```

## 2. Note that the species name given in "rownames(d)" and in "tre$tip.label" do not match.  Fix that (e.g. using str_replace() )

```{r}
rownames(d)
tre$tip.label

tre$tip.label <- str_replace(pattern = "\\.",
                             replacement =  "_",
                             string = tre$tip.label)
all.equal(rownames(d), tre$tip.label)
```

## 3. Should either range or size be transformed?

```{r}
d
summary(d)
# size in kg and range in km
```

Size definitely looks like it should be transformed given the large range and relatively small median size. I am going to transform both for now though

```{r}
dat_list <- list(
  N_spp = nrow(d),
  S = standardize(log(d$size)),
  R = standardize(log(d$range)),
  Imat = diag(nrow(d))
)

```


## 4. Does size predict range?  Evaluate this question without and with accounting for phylogenetic relationships.

### No phylogenic data


```{r}

no_phylo <- ulam(
  alist(
    R ~ multi_normal(mu , SIGMA),
    mu <- a + bS * S,
    matrix[N_spp, N_spp]:SIGMA <- Imat * sigma_sq,
    a ~ normal(0 , 1),
    bS ~ normal(0 , 0.5),
    sigma_sq ~ exponential(1)
  ),
  data = dat_list,
  chains = 4,
  cores = 4
)
precis(no_phylo)

```

### With phylo

```{r}
spp_obs <- rownames(d)
Brown <- corBrownian(phy = tre)
V <- vcv(Brown)
dat_list$V <- V[spp_obs, spp_obs]
dat_list$R <- dat_list$V / max(V)
Dmat <- cophenetic(tre)
dat_list$Dmat <- Dmat[spp_obs , spp_obs] / max(Dmat)

```

```{r}
# Brownian

phylo_Brown <- ulam(
  alist(
    R ~ multi_normal(mu, SIGMA),
    mu <- a + bS*S,
    matrix[N_spp,N_spp]: SIGMA <- R * sigma_sq,
    a ~ normal(0 , 1),
    c(bS) ~ normal(0 , 0.5),
    sigma_sq ~ exponential(1)
  ),
  data = dat_list ,
  chains = 4 ,
  cores = 4
)

precis(phylo_Brown)
```

```{r}
# OU

phylo_OU <- ulam(
  alist(
    R ~ multi_normal(mu, SIGMA),
    mu <- a + bS*S,
    matrix[N_spp,N_spp]: SIGMA <- cov_GPL1(Dmat , etasq , rhosq , 0.01),
    a ~ normal(0, 1),
    bS ~ normal(0, 0.5),
    etasq ~ half_normal(1, 0.25),
    rhosq ~ half_normal(3, 0.25)
  ),
  data = dat_list,
  chains = 4,
  cores = 4
)

precis(phylo_OU)
```

```{r}
dens(Dmat)
```


## 5. Discuss your findings.

Looks like adding phylogenetic data increased the effect of size on range in the OU model. I am currently concerned with how Dmat looks though. Something isn't right and I should probably fix it.